---
title: "Exercise 2 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

pacman::p_load(
  tidyverse, # Data manipulation,
  ggplot2, # beautiful figures,
  vdemdata, # download democracy datasets used in the scholarly literature.
  wbstats, # download data from Worldbank. Tremendous source of global socio-economic data.
  estimatr, # OLS with robust SE,
  modelsummary) # regression tables with nice layout,

# 1. The world Bank data
Dat <- wb_data("SI.POV.DDAY", # Download poverty data: <2.15$ per day,
               start_date = 2000, end_date = 2025) %>%
  rename(poverty = SI.POV.DDAY) %>% # rename poverty variable,
  select(country, date, poverty) %>% # Keep only 3 variables
  drop_na(poverty) %>% # Drop cases with missing data,
  group_by(country) %>% # Group by country,
  filter(date == max(date)) %>% # Keep the most recent data per country.
  mutate(date = as.numeric(date)) %>% ungroup()

# 2. Colonizer data
Dat <- Dat %>% 
  mutate(
    colonizer = case_when(
      str_detect(country, "Algeria|Benin|Burkina Faso|Cambodia|Central African Republic|Chad") |
        str_detect(country, "Djibouti|Gabon|Guinea|Laos|Haiti|Lebanon|Madagascar|Mali|Congo, Rep.|Cote D'Ivoire") | 
        str_detect(country, "Mauritania|Niger|Senegal|Syria|Togo|Tunisia|Viet Nam") ~ "France", # France
      str_detect(country, "Bahrain|Bangladesh|Botswana|Cyprus|Egypt|Fiji") |
        str_detect(country, "Gambia|Ghana|Guyana|India|Iraq|Jamaica|Jordan|Kenya|Kuwait") |
        str_detect(country, "Lesotho|Malawi|Malaysia|Mauritius|Myanmar|Nigeria|Oman") | 
        str_detect(country, "Pakistan|Qatar|Sierra Leone|Singapore") | 
        str_detect(country, "Sri Lanka|Sudan|Swaziland|Tanzania|Trinidad/Tobago|Uganda") | 
        str_detect(country, "United Arab Emirates|Yemen|Zambia|Zimbabwe") ~ "Britain", # Britain
      str_detect(country, "Burundi|Congo, Dem. Rep.|Rwanda") ~ "Belgium", # Belgium
      str_detect(country, "Iceland|Estonia") ~ "Denmark")) # Denmark

# 3. Independence from colonialism data
Dat <- Dat %>% 
  mutate(
    years_indep = case_when(
      is.na(colonizer) ~ as.numeric(NA),
      country == "Estonia" ~ date - 1645,
      country == "United States" ~ date - 1783, country == "Haiti" ~ date - 1804, 
      country == "Paraguay" ~ date - 1811, country == "Chile" ~ date - 1818,
      str_detect(country, "Argentina|Bolivia|Colombia") ~ date - 1819,
      str_detect(country, "Costa Rica|Dominican Republic|Mexico|Nicaragua|Panama|El Salvador|Guatemala|Honduras|Venezuela") ~ date - 1821,
      str_detect(country, "Brazil|Ecuador") ~ date - 1822,
      country == "Peru" ~ date - 1824, country == "Uruguay" ~ date - 1828,
      country == "Cuba" ~ date - 1899, country == "Australia" ~ date - 1901,
      country == "New Zealand" ~ date - 1907, country == "South Africa" ~ date - 1910,
      country == "Egypt" ~ date - 1922, country == "Iraq" ~ date - 1932,
      str_detect(country, "Iceland") ~ date - 1944,
      str_detect(country, "Korea|Taiwan|Vietnam") ~ date - 1945,
      str_detect(country, "Lebanon|Philippines|Syria") ~ date - 1946,
      str_detect(country, "Bangladesh|Pakistan|India|Liberia") ~ date - 1947,
      str_detect(country, "Myanmar|Israel|Jordan|Sri Lanka") ~ date - 1948,
      country == "Indonesia" ~ date - 1949, country == "Libya" ~ date - 1951,
      str_detect(country, "Cambodia|Loas") ~ date - 1954,
      str_detect(country, "Morocco|Sudan|Tunisia") ~ date - 1956,
      str_detect(country, "Malaysia|Ghana") ~ date - 1957,
      country == "Guinea" ~ date - 1958, country == "Singapore" ~ date - 1959,
      str_detect(country, "Benin|Burkina Faso|Central African Republic|Chad|Congo, Dem. Rep.|Congo, Rep.|Code D'Ivoire|Mali|Mauritania|Niger|Nigeria|Senegal|Gabon|Somalia|Togo") ~ date - 1960,
      str_detect(country, "Kuwait|Sierra Leone|Tanzania") ~ date - 1961,
      str_detect(country, "Algeria|Burundi|Rwanda|Jamaica|Trinidad/ Tobago|Uganda") ~ date - 1962,
      country == "Kenya" ~ date - 1963,
      str_detect(country, "Malawi|Zambia") ~ date - 1964,
      str_detect(country, "Gambia|Zimbabwe") ~ date - 1965,
      str_detect(country, "Botswana|Lesotho|Guyana") ~ date - 1966,
      str_detect(country, "Canada|Yemen") ~ date - 1967,
      str_detect(country, "Mauritius|Swaziland") ~ date - 1968,
      country == "Fiji" ~ date - 1970,
      str_detect(country, "Bahrain|Oman|Qatar|United Arab Emirates") ~ date - 1971,
      country == "Guinea-Bissau" ~ date - 1974,
      str_detect(country, "Angola|Mozambigue|Papua New Guinea") ~ date - 1975,
      country == "Djibouti" ~ date - 1977, 
      country == "Namibia" ~ date - 1990))

(Dat_vdem <- vdem %>% # The data are part of the package we loaded earlier
   as_tibble() %>% # turn into a tiblle
   select(country_name, country_text_id, year, v2xcl_rol) %>%
   rename(equal_liberty = v2xcl_rol, # rename liberty variable
          date = year,
          country = country_name)) #rename country variable

Dat <- inner_join(Dat, Dat_vdem, by = c("country", "date"))
```

1. Drop Haiti and Estonia from our dataset; they are Outliers.
```{r outliers, out.width='80%', fig.height = 4, fig.width = 5, results = FALSE}
Dat <- Dat %>%
  filter(country != "Haiti" & country != "Estonia")
```

2. Visualize a prediction from a regression: How would citizenship rights look on average in the range of 5 years before independence to 50 years after independence from a colonizer? Add a red line that marks independence.
`r hide("R solution -> dont' peak too early ;) !")`
```{r plot1, out.width='80%', fig.height = 4, fig.width = 5, results = FALSE}
ols <- lm_robust(equal_liberty ~  years_indep, data = Dat)

(fict_dat <- tibble(years_indep = -5:50))

(fict_dat <- predict( #<<
  ols, newdata = fict_dat, #<<
  interval = "confidence", level = 0.95)$fit %>% #<<
   as_tibble() %>% # Turn into a tibble, then
   bind_cols(fict_dat, .)) # Add to the synthetic data.

ggplot(data = fict_dat, aes(y = fit, x = years_indep)) +
  geom_vline(xintercept = 0, color = "red", lty = "dashed") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.5) +
  geom_line() +
  labs(
    title = "Predictions based on regression of citizenship rights",
    x = "By years since independence",
    # Write Greek beta into axis title.
    y = "Predicted average of (Freedomhouse) citizenship rights") +
  theme_minimal()
```
`r unhide()`

3. Make a coefficient plot that shows how both Denmark and the other former colonizers fare compared to Belgium.
`r hide("R solution -> dont' peak too early ;) !")`
```{r plot2, out.width='80%', fig.height = 4, fig.width = 5, results = FALSE}
(plotdata <- lm_robust(equal_liberty ~ colonizer, data = Dat) %>%
   tidy() %>% # Turn results into a tibble, #<<
   mutate( # Rename variables for the plot.
     term = case_when(
       term == "colonizerFrance" ~ "France",
       term == "colonizerBritain" ~ "Britain",
       term == "colonizerDenmark" ~ "Denmark",
       term == "(Intercept)" ~ "Intercept \n (Belgium)")) %>%
  filter(term != "Intercept \n (Belgium)"))

ggplot(data = plotdata, aes(y = estimate, 
                            # Order by effect size
                            x = reorder(term, estimate))) +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +
  # Point with error-bars,
  geom_pointrange(aes(min = conf.low, max = conf.high)) +
  coord_flip() + # Flip Y- & X-Axis,
  labs(
    title = "Regression of citizen rights",
    x = "Colonizer",
    # Write Greek beta into axis title.
    y = expression("Estimate of"~beta~"(Difference to countries colonized by Belium)")) +
  theme_minimal()
```
`r unhide()`