---
title: "Exercise 3 (15 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

```{r results = FALSE, echo = FALSE}
pacman::p_load(
  tidyverse, # Data manipulation,
  haven, # Handle labelled data.
  ggplot2, # beautiful figures,
  estimatr, # Regression for weighted data,
  modelr, 
  texreg) # regression tables with nice layout.

# 1. read the ESS round 9 data
ESS <- read_spss("../../../assets/ESS9e03_1.sav") %>%
  filter(cntry == "DK") %>% # Keep only the Danish data,
  # Keep only a minimum set of variables we need today,
  select(idno, pspwght, gndr, eduyrs, agea, 
         psppsgva, trstlgl, trstplc) %>%
  mutate(psppsgva = zap_labels(psppsgva), # Make numeric
         trstlgl = zap_labels(trstlgl), # Make numeric
         trstplc = zap_labels(trstplc), # Make numeric
         pspwght = zap_labels(pspwght), # Make numeric
         eduyrs = case_when( # Censor years of education at 9 & 21 years.
           eduyrs > 21 ~ 21,
           eduyrs < 9 ~ 9,
           TRUE ~ as.numeric(eduyrs)),
         gndr = as_factor(gndr)) %>% # Make factor
drop_na() # Delete cases with missing values.
```

1. Regress trust in the legal system on years of education. Use the `sumary()` command to display the output. 

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
# Run the regression and make a regression table.
mod1 <- lm_robust(trstlgl ~ eduyrs, data = ESS, weights = pspwght)
summary(mod1)
```
`r unhide()`
```{r include = FALSE}
# Run the regression and make a regression table.
results <- tidy(mod1)
teh <- results$statistic[results$term == "eduyrs"] %>% as.scalar()
p <- results$p.value[results$term == "eduyrs"] %>% as.scalar()
```

2. What does the $t$-value of `teh` indicate? `r mcq(c("It's too large to be significant", "Its too small to be significant", answer = "The slope estimate beta is that many standard errors away from 0", "The probability that our hypothesis is correct is only ca 5%"))`.

3. What is the correct interpretation of the *p*-value `p`? `r mcq(c("The probability that education predicts trust is very low", "The p-value is too small to be significant", "The probability of the H0 is very small", answer = "The probability of randomly drawing a sample with this slope estimate, given that the slope estimate is 0 in reality"))`.

4. Can we reject the $H_{0}$?

5. Is the association between trust in the legal system and education statistically significant?

