---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

1. Download "Legewie_ESS_02.dta" from Absalon. Run the code below to prepare the data for analysis.
```{r results = FALSE}
pacman::p_load(
  tidyverse, # Data manipulation,
  haven, # Handle Stata data.
  ggplot2, # beautiful figures,
  estimatr, # Regression for weighted data,
  furniture, # For row-means,
  texreg) # regression tables with nice layout.

# Bali terror attack
event_date_begin <- as.Date("2002-09-14")
event_date       <- as.Date("2002-10-13")
event_date_end   <- as.Date("2002-10-20")

# Read the ESS 2002 data and prepare for analysis
ESS <- read_dta("../../../assets/Legewie_ESS_02.dta") %>%
  mutate(
    # Day and time of the survey interview,
    int_year = inwyr,
    int_mon  = inwmm,
    int_day  = inwdd,
    int_date = sprintf("%s-%s-%s", int_year, int_mon, int_day) %>% as.Date(),
    # Treatment indicator,
    treat = case_when(
      int_date > event_date & int_date <= event_date_end ~ "After the attack",
      int_date < event_date & int_date > event_date_begin ~ "Before the attack",
      TRUE ~ as.character(NA)) %>% fct_relevel("Before the attack", "After the attack"),
    # Attitudes towards immigrants
    anti_immi = rowmeans(
      imtcjob, imbleco, imbgeco, imueclt,
      imwbcnt, imwbcrm, imbghct, na.rm = TRUE) %>% scale() %>% as.numeric(),
    # Subtract from max to make larger values more xenophobic.
    anti_immi = max(anti_immi, na.rm = TRUE) - anti_immi,
    # Immigrant origin
    brncntr = as_factor(brncntr),
    mocntr = as_factor(mocntr),
    facntr = as_factor(facntr),
    # Weights
    pspwght = pweight * dweight) %>%
  # Sample restriction        
  filter(str_detect(cntry, "PT") & # Countries
           # Mainstream members (i.e. immigrant origin)
           brncntr == "yes" & mocntr == "yes" &  facntr == "yes") %>%
  # Keep only a minimum set of variables we need today,
  select(treat, anti_immi, cntry, int_date, pspwght) %>%
  drop_na() # Delete cases with missing values.
```

```{r olsPT, include = FALSE}
# Estimate the treatment effect with OLS
ols_pt <- lm_robust(anti_immi ~ treat, data = ESS, weights = pspwght)

# Report results in a nicely-formatted table
screenreg(ols_pt, include.ci = FALSE, digits = 3)
```

```{r echo = FALSE}
coefPT <- coef(ols_pt)["treatAfter the attack"] %>% round(., 3) %>% as.scalar()
se <- vcov(ols_pt) %>% diag()
se <- se["treatAfter the attack"] %>% sqrt() %>% round(., 3) %>% as.scalar()
```

2. What is the average causal effect of the Bali 2002 terror attack among the Portuguese? It is `r fitb(coefPT, width = "5")` standard deviations.
`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label = "olsPT"}
```
`r unhide()`

3. Is the effects statistically significant? Yes, the standard error is `r fitb(se, width = "5")` and thus less than half the size of the estimated causal effect. This implies the *t*-value is larger than the conventional theshold of `r fitb(1.96, width = "4")`.

4. Make a coefficient plot of the result; don't report the intercept.
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
plotdata_PT <- ols_pt %>%
  tidy() %>% # Turn into tibble
  filter(term != "(Intercept)") %>% # Delete intercept from data
  mutate(
    term = case_when(
      term == "treatAfter the attack" ~ "After the attack"))

ggplot(data = plotdata_PT, aes(y = estimate, x = term, ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") + 
  geom_pointrange() + 
  coord_flip() +
  theme_minimal() + 
  labs(title = "Causal effect of Bali 2002 terror attack",
    x = "", y =  "Difference in xenophobia compared to before the attack
    (in standard deviations)")
```
`r unhide()`

```{r olsSE, include = FALSE}
# Read the ESS 2002 data and prepare for analysis
ESS <- read_dta("../../../assets/Legewie_ESS_02.dta") %>%
  mutate(
    # Day and time of the survey interview,
    int_year = inwyr,
    int_mon  = inwmm,
    int_day  = inwdd,
    int_date = sprintf("%s-%s-%s", int_year, int_mon, int_day) %>% as.Date(),
    # Treatment indicator,
    treat = case_when(
      int_date > event_date & int_date <= event_date_end ~ "After the attack",
      int_date < event_date & int_date > event_date_begin ~ "Before the attack",
      TRUE ~ as.character(NA)) %>% fct_relevel("Before the attack", "After the attack"),
    # Attitudes towards immigrants
    anti_immi = rowmeans(
      imtcjob, imbleco, imbgeco, imueclt,
      imwbcnt, imwbcrm, imbghct, na.rm = TRUE) %>% scale() %>% as.numeric(),
    # Subtract from max to make larger values more xenophobic.
    anti_immi = max(anti_immi, na.rm = TRUE) - anti_immi,
    # Immigrant origin
    brncntr = as_factor(brncntr),
    mocntr = as_factor(mocntr),
    facntr = as_factor(facntr),
    # Weights
    pspwght = pweight * dweight) %>%
  # Sample restriction        
  filter(str_detect(cntry, "PT|SE") & # Countries
           # Mainstream members (i.e. immigrant origin)
           brncntr == "yes" & mocntr == "yes" &  facntr == "yes") %>%
  # Keep only a minimum set of variables we need today,
  select(treat, anti_immi, cntry, int_date, pspwght) %>%
  drop_na() # Delete cases with missing values.

# Estimate the treatment effect with OLS
ols_se <- lm_robust(anti_immi ~ treat, 
                    data = ESS %>% filter(cntry == "SE"), 
                    weights = pspwght)

# Report results in a nicely-formatted table
screenreg(ols_se, include.ci = FALSE, digits = 3)
```

```{r echo = FALSE}
coefSE <- coef(ols_se)["treatAfter the attack"] %>% round(., 3) %>% as.scalar()
```

4. Adapt the data preparation code so that the sample also entails Swedes in addition to Portuguese respondents. The causal effect of the Bali 2002 terror attack among Swedes is `r fitb(coefSE, width = "5")` and `r mcq(c("Significant", "Statistically significant", answer = "Not statistically significant"))`.
`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label = "olsSE"}
```
`r unhide()`

5. Generate "plotdata" for the Swedish and Portuguese results. Use `bind_rows(plotdata_SE, plotdata_PT, .id = "cntry")` to combine both into one tibble. Then make a coefficient plot of both effects. *Look at you, you have learned to conduct analysis like sociologists who publish in the American Journal of Sociology!*
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
plotdata_SE <- ols_se %>%
  tidy() %>% # Turn into tibble
  filter(term != "(Intercept)") %>% # Delete intercept from data
  mutate(
    term = case_when(
      term == "treatAfter the attack" ~ "After the attack"))

plotdata <- bind_rows(plotdata_SE, plotdata_PT, .id = "cntry") %>%
  mutate(
    cntry = case_when(
      cntry == 1 ~ "Sweden",
      cntry == 2 ~ "Portugal"))

ggplot(data = plotdata, aes(y = estimate, x = reorder(cntry, estimate), 
                            ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") + 
  geom_pointrange() + 
  coord_flip() +
  theme_minimal() + 
  labs(title = "Causal effect of Bali 2002 terror attack",
    x = "", y =  "Difference in xenophobia compared to before the attack
    (in standard deviations)")
```
`r unhide()`