---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

1. We take another look at "Legewie_ESS_02.dta". Adapt the preparation of the data from last week: Add the following variables to the data: Age (in years, based on `int_year` and `yrbrn`), gender (based on `gndr`), employment status (with categories "employed", "unemployed", "retired", "other",  based on `pdwrk`, `uempla`, and `rtrd`). Also keep data for Finnland, Sweden, and Portugal.
`r hide("R solution -> dont' peak to early ;) !")`
```{r results = FALSE}
pacman::p_load(
  tidyverse, # Data manipulation,
  haven, # Handle Stata data.
  ggplot2, # beautiful figures,
  estimatr, # Regression for weighted data,
  furniture, # For row-means,
  texreg) # regression tables with nice layout.

# Bali terror attack
event_date_begin <- as.Date("2002-09-14")
event_date       <- as.Date("2002-10-13")
event_date_end   <- as.Date("2002-10-20")

# Read the ESS 2002 data and prepare for analysis
ESS <- read_dta("../../../assets/Legewie_ESS_02.dta") %>%
  mutate(
    # Day and time of the survey interview,
    int_year = inwyr,
    int_mon  = inwmm,
    int_day  = inwdd,
    int_date = sprintf("%s-%s-%s", int_year, int_mon, int_day) %>% as.Date(),
    # Treatment indicator,
    treat = case_when(
      int_date > event_date & int_date <= event_date_end ~ "After the attack",
      int_date < event_date & int_date > event_date_begin ~ "Before the attack",
      TRUE ~ as.character(NA)) %>% fct_relevel("Before the attack", "After the attack"),
    # Attitudes towards immigrants
    anti_immi = rowmeans(
      imtcjob, imbleco, imbgeco, imueclt,
      imwbcnt, imwbcrm, imbghct, na.rm = TRUE) %>% scale() %>% as.numeric(),
    # Subtract from max to make larger values more xenophobic.
    anti_immi = max(anti_immi, na.rm = TRUE) - anti_immi,
    # Immigrant origin
    brncntr = as_factor(brncntr),
    mocntr = as_factor(mocntr),
    facntr = as_factor(facntr),
    #######################################
    # Age
    age = int_year - yrbrn, #<<
    # Gender
    gndr = as_factor(gndr) %>% fct_drop(),
    # Employment status
    empl_stat = case_when(
      pdwrk == 1 ~ "Working",
      uempla == 1 ~ "Unemployed",
      rtrd == 1 ~ "Retired",
      TRUE ~ "Other"),
    ########################################
    # Weights
    pspwght = pweight * dweight) %>%
  # Sample restriction        
  filter(str_detect(cntry, "PT|SE|FI") & # Countries
           # Mainstream members (i.e. immigrant origin)
           brncntr == "yes" & mocntr == "yes" &  facntr == "yes") %>%
  # Keep only a minimum set of variables we need today,
  select(treat, anti_immi, cntry, int_date, pspwght,
         age, gndr, eduyrs, empl_stat) %>%
  drop_na() # Delete cases with missing values.
```
`r unhide()`


```{r olsPT, include = FALSE}
# Estimate the controlled treatment effect with multiple OLS
ols_pt_adj <- lm_robust(anti_immi ~ treat + age + gndr + empl_stat, 
                        data = ESS %>% filter(cntry == "PT"), 
                        weights = pspwght)

# Report results in a nicely-formatted table
screenreg(ols_pt_adj, 
          include.ci = FALSE, digits = 3)
```

```{r echo = FALSE}
coefPT <- coef(ols_pt_adj)["treatAfter the attack"] %>% round(., 3) %>% as.scalar()
se <- vcov(ols_pt_adj) %>% diag()
se <- se["treatAfter the attack"] %>% sqrt() %>% round(., 3) %>% as.scalar()
```

2. What is the average causal effect of the Bali 2002 terror attack among the Portuguese, controlled for age, gender, years of education, and employment status? It is `r fitb(coefPT, width = "5")` standard deviations.
`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label = "olsPT"}
```
`r unhide()`

3. Is the effects statistically significant? Yes, the standard error is `r fitb(se, width = "5")` and thus less than half the size of the estimated causal effect. This implies the *t*-value is larger than the conventional threshold of `r fitb(1.96, width = "4")`.

4. Also estimate the bivariate unadjusted effect and make a coefficient plot of the two results; don't report the intercept. Use `bind_rows(..., .id = "controlled")` to combine both into one tibble. What do we learn by comparing the two estimates? `r mcq(c("Bot much, both effects are significant.", answer = "The bivariate effect is a bit too large; there was some selection into treatment that introduced omitted variable bias", "The result from multiple OLS is more uncertain and hence less trustworthy"))`.

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
# Estimate the bivariate treatment effect with OLS
ols_pt_biv <- lm_robust(anti_immi ~ treat, 
                        data = ESS %>% filter(cntry == "PT"), 
                        weights = pspwght)

# Add estimation results in to tibble
plotdata <- bind_rows(
  tidy(ols_pt_biv), 
  tidy(ols_pt_adj), 
  .id = "adjusted") %>%
  mutate(
    adjusted = case_when(
      adjusted == 1 ~ "Bivariate",
      adjusted == 2 ~ "Adjusted")) %>%
  filter(term == "treatAfter the attack")

ggplot(data = plotdata, aes(y = estimate, x = adjusted, ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") + 
  geom_pointrange() + 
  coord_flip() +
  theme_minimal() + 
  labs(title = "Causal effect of Bali 2002 terror attack among the Portuguese",
    x = "", y =  "Difference in xenophobia compared to before the attack
    (in standard deviations)")
```
`r unhide()`

```{r olsSE, include = FALSE}
# Estimate the controlled treatment effect with multiple OLS
ols_fi_adj <- lm_robust(anti_immi ~ treat + age + gndr + empl_stat, 
                        data = ESS %>% filter(cntry == "FI"), 
                        weights = pspwght)

# Report results in a nicely-formatted table
screenreg(ols_fi_adj, 
          include.ci = FALSE, digits = 3)
```

```{r echo = FALSE}
coefFI <- coef(ols_fi_adj)["treatAfter the attack"] %>% round(., 3) %>% as.scalar()
```

5. The adjusted causal effect of the Bali 2002 terror attack among Finns is `r fitb(coefFI, width = "5")` and `r mcq(c("Significant", "Statistically significant", answer = "Not statistically significant"))`.
`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label = "olsSE"}
```
`r unhide()`