---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
```

1. Do the [Setup](https://merlin-ols.netlify.app/about/), if you have not yet done so. 

2. Start a new R-script and copy the lines of R code from my slides that generate 'Dat'. That is, get Freedom House and WB data via the APIs, join them, and construct our socialism index.
  
`r hide("R solution -> dont' peak too early ;) !")`
```{r}
pacman::p_load(
  tidyverse, # Data manipulation,
  ggplot2, # beautiful figures,
  wbstats, # download data from Worldbank. Tremendous source of global socio-economic data.
  estimatr, # OLS with robust SE,
  democracyData, # download democracy data used in the scholarly literature.
  countrycode, # Easy recodings of country names,
  remotes) # Install beta version packages from GitHub.

# 1. Download Freedom House data for all countries since 1972,
Dat_citi_rights <- download_fh(verbose = FALSE) %>% 
   rename(country = fh_country, # rename country ID,
          citizen_rights = fh_total_reversed, # rename Citizenship rights indicator,
          date = year) %>% # rename year,
   select(country, date, citizen_rights) # Keep only these 3 variables.

# 2. Download poverty data: <US$2.15 per day
Dat_poverty <- wb_data("SI.POV.DDAY", 
                        start_date = 2002, end_date = 2022) %>% #<<
   rename(poverty = SI.POV.DDAY) %>% # rename poverty variable,
   select(country, date, poverty) %>% # Keep only 3 variables
   drop_na(poverty) %>% # Drop cases with missing data,
   group_by(country) %>% # Group by country,
   filter(date == max(date)) %>% # Keep the most recent data per country.
   ungroup()

# 3. Join Freedom House and Worldbank data
Dat <- inner_join(Dat_poverty, Dat_citi_rights, by = c("country", "date"))

# 4. Code socialism index
Dat <- Dat %>% mutate(socialist = case_when( # Years socialist minus years since,
  country == "China" ~ date - 1949,
  country == "Vietnam" ~ date - 1945,
  country == "Algeria" ~ date - 1962,
  str_detect(country,"Portugal|Bangladesh") ~ date - 1972,
  country == "Guinea-Bissau" ~ date - 1973,
  country == "India" ~ date - 1976,
  country == "Nicaragua" ~ date - 1979,
  country == "Sri Lanka" ~ date - 1978,
  country == "Tanzania" ~ date - 1964,
  country == "Albania" ~ (1992 - 1944) - (date - 1992),
  str_detect(country, "Angola|Cabo Verde|Madagascar") ~ (1992 - 1975) - (date - 1992),
  str_detect(country,"Belarus|Bulgaria") ~ (1990 - 1946) - (date - 1990),
  str_detect(country, "Benin|Mozambique") ~ (1990 - 1975) - (date - 1990),
  country == "Chad" ~ (1975 - 1962) - (date - 1975), country == "Congo, Rep." ~ (1992 - 1970) - (date - 1992),
  country == "Czech Republic" ~ (1990 - 1948) - (date - 1990), country == "Djibouti" ~ (1992 - 1981) - (date - 1992),
  country == "Ethiopia" ~ (1991 - 1974) - (date - 1991), country == "Ghana" ~ (1966 - 1960) - (date - 1966),
  country == "Guinea" ~ (1984 - 1958) - (date - 1984), country == "Hungary" ~ (1989 - 1949) - (date - 1989),
  country == "Iraq" ~ (2005 - 1958) - (date - 2005), country == "Mali" ~ (1991 - 1960) - (date - 1991),
  country == "Mauritania" ~ (1978 - 1961) - (date - 1978), country == "Mongolia" ~ (1992 - 1924) - (date - 1992),
  country == "Myanmar" ~ (1988 - 1962) - (date - 1988), country == "Poland" ~ (1989 - 1945) - (date - 1989),
  country == "Romania" ~ (1989 - 1947) - (date - 1989), country == "Russian Federation" ~ (1991 - 1922) - (date - 1991),
  country == "Seychelles" ~ (1991 - 1977) - (date - 1991), country == "Senegal" ~ (1981 - 1960) - (date - 1981),
  country == "Sierra Leone" ~ (1991 - 1978) - (date - 1991), country == "Somalia" ~ (1991 - 1969) - (date - 1991),
  country == "Sudan" ~ (1985 - 1969) - (date - 1985), country == "Syria" ~ (2012 - 1963) - (date - 2012),
  country == "Tunisia" ~ (1988 - 1964) - (date - 1988), country == "Ukraine" ~ (1991 - 1919) - (date - 1991),
  country == "Yemen, Rep." ~ (1991 - 1967) - (date - 1991), country == "Zambia" ~ (1991 - 1973) - (date - 1991),
  str_detect(country,"Slovenia|Croatia|Serbia|Montenegro|Bosnia and Herzegovina|North Macedonia|Kosovo") ~ (1992 - 1943) - (date - 1992),
  TRUE ~ 0),
  socialist = case_when( # Min. 5 years given socialist history,
    (socialist < 5 & socialist > 0) | socialist < 0 ~ 5,
    TRUE ~ socialist)) %>% drop_na() # Drop countries with missing values.
```
`r unhide()`

3. Make a scatterplot with poverty as outcome on the Y-axis and civic and political citizenship rights as predictor on the X-axis.
`r hide("R solution -> dont' peak too early ;) !")`
```{r out.width='100%', fig.height = 4, fig.width = 7.5, results = FALSE, echo = TRUE}
ggplot(data = Dat, aes(y = poverty, x = citizen_rights)) +
  geom_text(aes(label = country)) +
  labs(y = "% population with less than US$2.15 a day\n (less than DKK420 per month)", 
       x = "Freedom House index of citizenship rights") +
  theme_minimal() +
  guides(color = "none")
```
`r unhide()`

```{r results = FALSE, echo = FALSE}
cor_povcitiz <- cor_povcitiz <- Dat %>% 
  select(poverty, citizen_rights) %>% 
  cor() %>% 
  round(., 2)
cor_povcitiz <- cor_povcitiz[1, 2]
```

4. Estimate the correlation between poverty and civic and political citizenship rights (rounded to two digits): `r fitb(cor_povcitiz, width = "5")`
`r hide("R solution -> dont' peak too early ;) !")`
```{r}
Dat %>% 
  select(poverty, citizen_rights) %>% 
  cor() %>% 
  round(., 2)
```
`r unhide()`

4. What does the correlation suggest about whether there is a freedom-equality trade off?
`r mcq(c(answer = "No, where people have civic and political rights, there is also less poverty", "Yes, limited freedom is the price to pay for equality", "The result is inconclusive"))`
