---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
```

```{r results = FALSE, echo = FALSE}
pacman::p_load(
  tidyverse, # Data manipulation,
  haven, # Handle labelled data.
  ggplot2, # beautiful figures,
  estimatr, # Regression for weighted data,
  essentials,
  modelsummary) # regression tables with nice layout.

# 1. read the ESS round 9 data
ESS <- read_spss("../../../assets/ESS9e03_1.sav") %>%
  filter(cntry == "DK") %>% # Keep only the Danish data,
  # Keep only a minimum set of variables we need today,
  select(idno, pspwght, gndr, eduyrs, agea, 
         psppsgva, trstlgl, trstplc) %>%
  mutate(psppsgva = zap_labels(psppsgva), # Make numeric
         trstlgl = zap_labels(trstlgl), # Make numeric
         trstplc = zap_labels(trstplc), # Make numeric
         pspwght = zap_labels(pspwght), # Make numeric
         eduyrs = case_when( # Censor years of education at 9 & 21 years.
           eduyrs > 21 ~ 21,
           eduyrs < 9 ~ 9,
           TRUE ~ as.numeric(eduyrs)),
         gndr = as_factor(gndr)) %>% # Make factor
drop_na() # Delete cases with missing values.

# Run the regression and make a regression table.
mod1 <- lm_robust(trstplc ~ eduyrs, data = ESS)
coef <- coef(mod1)["eduyrs"] %>% round(., 3) %>% as.scalar()
```

1. Use the ESS with exactly the variables I use for today. Disregarding the weights for a moment: How much is a one-year increase in education associated with an average change in trust in the police? The answer is `r fitb(coef, width = "5")`
  
`r hide("R solution -> dont' peak too early ;) !")`
```{r}
pacman::p_load(
  tidyverse, # Data manipulation,
  haven, # Handle labelled data.
  ggplot2, # beautiful figures,
  estimatr, # Regression for weighted data,
  modelsummary) # regression tables with nice layout.

# 1. read the ESS round 9 data
ESS <- read_spss("../../../assets/ESS9e03_1.sav") %>%
  filter(cntry == "DK") %>% # Keep only the Danish data,
  # Keep only a minimum set of variables we need today,
  select(idno, pspwght, gndr, eduyrs, agea, 
         psppsgva, trstlgl, trstplc) %>%
  mutate(psppsgva = zap_labels(psppsgva), # Make numeric
         trstlgl = zap_labels(trstlgl), # Make numeric
         trstplc = zap_labels(trstplc), # Make numeric
         pspwght = zap_labels(pspwght), # Make numeric
         eduyrs = case_when( # Censor years of education at 9 & 21 years.
           eduyrs > 21 ~ 21,
           eduyrs < 9 ~ 9,
           TRUE ~ as.numeric(eduyrs)),
         gndr = as_factor(gndr)) %>% # Make factor
drop_na() # Delete cases with missing values.

# Run the regression and make a regression table.
mod1 <- lm_robust(trstplc ~ eduyrs, data = ESS)
modelsummary(list("Tr. police" = mod1), statistic = NULL, gof_map = c("nobs", "r.squared"))
```
`r unhide()`

2. Still ignoring the weights: Is there a gender-difference in trust in the police? `r mcq(c(answer = "Yes, women seem to trust the police a bit more.", "No, women and men trust the police exactly the same", "Yes, men are more trusting of the police"))`.
`r hide("R solution -> dont' peak too early ;) !")`
```{r}
# Run the regression and make a regression table.
mod2 <- lm_robust(trstplc ~ gndr, data = ESS)
modelsummary(list("Tr. police" = mod2), statistic = NULL, gof_map = c("nobs", "r.squared"))
```
`r unhide()`

3. Rerun both analyses using weights. Report all result in one regression table. Do the weights matter? `r mcq(c("Yes, women are now less trusting", "Yes, education is now a stronger predictor", answer = "Yes, now women trust more and education becomes less important", "No, the reuslts don't change much"))`.

`r hide("R solution -> dont' peak too early ;) !")`
```{r}
# Run the regression and make a regression table.
mod1_weighted <- lm_robust(trstplc ~ eduyrs, data = ESS, weights = pspwght)
mod2_weighted <- lm_robust(trstplc ~ gndr, data = ESS, weights = pspwght)
# Make one regression table.
modelsummary(list("Unweighted" = mod1, "Unweighted" = mod2, 
                  "Weighted" = mod1_weighted, "Weighted" = mod2_weighted),
             statistic = NULL, gof_map = c("nobs", "r.squared"))
```
`r unhide()`
