---
title: "Exercise 1 (25 minutes)"
subtitle: "You practice: a full IV workflow on a new instrument — first stage, reduced form, Wald ratio, then 2SLS with controls. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 3)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

num_fitb <- function(x, digits = 3, width = 7) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}
```

```{r hidden-setup, include = FALSE}
pacman::p_load(tidyverse, estimatr, ivreg, modelsummary, masteringmetrics)
data("child_labor", package = "masteringmetrics")
```

**Acemoglu & Angrist (2000) had an idea.** US federal states required children to have a minimum
amount of schooling before they were allowed to work — and *how much* differed across states and
changed over time. They argued that the exact year a given state reformed its child-labour law was
essentially arbitrary. That gives us an instrument for schooling.

```{tikz dag-ex1, echo = FALSE, out.width='42%'}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning}
\definecolor{kured}{HTML}{901A1E}
\begin{tikzpicture}[>=Latex, semithick]
\sffamily
\node[ellipse, draw=kured, thick, align=center] (Z) at (0,0) {Child-labour\\law};
\node[ellipse, draw] (D) at (3.4,0) {Education};
\node[ellipse, draw=gray, dashed, align=center] (C) at (3.4,2.2) {Ability, family\\background};
\node[ellipse, draw] (Y) at (7,0) {Earnings};
\draw[->, kured, thick] (Z) -- (D);
\draw[->] (D) -- (Y);
\draw[->, gray, dashed] (C) -- (D);
\draw[->, gray, dashed] (C) -- (Y);
\end{tikzpicture}
```

1. Load the data and build the instrument. `cl7`, `cl8` and `cl9` are dummies for living in a state
   that required 7, 8 or 9 years of schooling. Collapse them into **one continuous** variable, and
   make state of birth and census year **categorical**.

```{r prep}
pacman::p_load(tidyverse, estimatr, ivreg, modelsummary, masteringmetrics)

data("child_labor", package = "masteringmetrics")

child_labor <- child_labor %>%
  mutate(
    # One continuous child-labour-law variable: 7, 8 or 9 required years
    chld_lb_lw = case_when(
      cl7 == 1 ~ 7,
      cl8 == 1 ~ 8,
      cl9 == 1 ~ 9
    ),
    sob  = factor(sob),  # US state of birth
    year = factor(year)  # Census year
  ) %>%
  drop_na(chld_lb_lw)
```

`r hide("Stuck?")`
- Did `pacman::p_load(...)` run without red errors? `ivreg` is new this week — if it is missing,
  run `install.packages("ivreg")` once.
- `data()` can't find `child_labor`? It ships inside **`masteringmetrics`**, which is a GitHub
  package — see **About & setup** on the course website.
- This data set has ~650,000 rows, so models take a few seconds. That is normal, not a crash.
- Still stuck? Ask in the [t-R-ouble forum](https://absalon.ku.dk/courses/70545/discussion_topics/563046?module_item_id=2037268) on Absalon.

`r unhide()`

```{r solutions, include = FALSE}
fs <- lm_robust(indEduc ~ chld_lb_lw + yob + year + sob, data = child_labor)
rf <- lm_robust(lnwkwage ~ chld_lb_lw + yob + year + sob, data = child_labor)
phi <- coef(fs)["chld_lb_lw"]
rho <- coef(rf)["chld_lb_lw"]
wald <- rho / phi
iv <- ivreg(lnwkwage ~ indEduc + yob + year + sob | chld_lb_lw + yob + year + sob,
            data = child_labor)
b_iv <- coef(iv)["indEduc"]
```

2. **Predict before you compute.** The instrument is "how many years of schooling the law required".
   Before running anything: which sign do you expect for the **first stage** — the effect of the law
   on actual years of education?
   `r mcq(c(answer = "Positive — a stricter law should keep children in school longer", "Negative — stricter laws push children out of school", "Zero — laws are rarely enforced"))`

3. **Estimate the first stage.** Regress `indEduc` on the instrument, controlling for year of birth
   (`yob`), census year (`year`) and state of birth (`sob`).

`r hide("Hint")`
`lm_robust(indEduc ~ chld_lb_lw + yob + year + sob, data = child_labor)`
`r unhide()`

`r hide("R solution — don't peek too early ;) !")`
```{r sol3, eval = FALSE}
first_stage <- lm_robust(indEduc ~ chld_lb_lw + yob + year + sob,
                         data = child_labor)

modelsummary(list("Years of education" = first_stage),
             coef_map = c("chld_lb_lw" = "Required years of schooling"),
             stars = TRUE, gof_map = c("nobs"), output = "kableExtra")
```
`r unhide()`

   One extra year of *required* schooling translates into `r num_fitb(phi)` extra years of
   *actual* education. Call this $\phi$.

4. **Estimate the reduced form.** `lnwkwage` is the log weekly wage. Use the **same** controls.

`r hide("R solution — don't peek too early ;) !")`
```{r sol4, eval = FALSE}
reduced_form <- lm_robust(lnwkwage ~ chld_lb_lw + yob + year + sob,
                          data = child_labor)

modelsummary(list("log weekly wage" = reduced_form),
             coef_map = c("chld_lb_lw" = "Required years of schooling"),
             stars = TRUE, gof_map = c("nobs"), output = "kableExtra")
```
`r unhide()`

   One extra required year raises log wages by `r num_fitb(rho, digits = 4)`. Call this $\rho$.
   This number is the causal effect **of the law** — *not* of education —
   `r mcq(c(answer = "provided the timing of state law reforms really is as-if random", "always, because it is a regression", "only if education is randomly assigned"))`.

5. **Compute the Wald ratio by hand.** Divide, and mind which goes on top.

   $$\lambda = \frac{\rho}{\phi} = \;?$$

   Your answer: `r num_fitb(wald)`

`r hide("R solution")`
```{r sol5, eval = FALSE}
phi <- coef(first_stage)["chld_lb_lw"]
rho <- coef(reduced_form)["chld_lb_lw"]

rho / phi
```
`r unhide()`

6. **Now do it properly with `ivreg()`.** Write the 2SLS in one line. Remember: the controls must
   appear on **both** sides of the `|`.

`r hide("Hint")`
The pattern is `ivreg(outcome ~ treatment + controls | instrument + controls, data = ...)`.
Here the treatment is `indEduc` and the instrument is `chld_lb_lw`.
`r unhide()`

`r hide("R solution — don't peek too early ;) !")`
```{r sol6, eval = FALSE}
twosls <- ivreg(
  lnwkwage ~ indEduc + yob + year + sob | chld_lb_lw + yob + year + sob,
  data = child_labor
)

modelsummary(list("2SLS" = twosls),
             coef_map = c("indEduc" = "Years of education"),
             stars = TRUE, gof_map = c("nobs"), output = "kableExtra")
```
`r unhide()`

   The 2SLS estimate is `r num_fitb(b_iv)` — compared with your hand-computed Wald ratio of
   `r round(wald, 3)`. They are
   `r mcq(c(answer = "identical — 2SLS reproduces the Wald ratio", "close but meaningfully different", "unrelated quantities"))`.

7. **Write one sentence** as a `# comment` interpreting the 2SLS estimate in plain language, with
   units — then compare:

`r hide("Model answer")`
Among men whose schooling was actually changed by child-labour laws, one additional year of
education raises weekly wages by about **14%**. Because the estimate uses only the variation in
schooling driven by the timing of state law reforms, it is not contaminated by ability or family
background — *provided* that timing really was as-if random and affected wages only through
schooling. Note this is a **LATE**: it describes compliers — children who would have left school
earlier under a laxer law — not the population as a whole.
`r unhide()`

8. **Bonus, for the fast.** Compare the 2SLS estimate to a naïve OLS of `lnwkwage` on `indEduc`
   with the same controls. Which is larger, and does that match the ability-bias prediction?

`r hide("R solution")`
```{r sol8, eval = FALSE}
ols <- lm_robust(lnwkwage ~ indEduc + yob + year + sob, data = child_labor)

modelsummary(list("OLS" = ols, "2SLS" = twosls),
             coef_map = c("indEduc" = "Years of education"),
             stars = TRUE, gof_map = c("nobs"), output = "kableExtra")
```

As in the lecture, the IV estimate comes out **larger** than OLS — the opposite of what simple
ability bias predicts. The same three explanations apply: LATE versus ATE (compliers are people
pushed through school by law, whose returns may be higher), measurement error in schooling biasing
OLS toward zero, and instrument weakness.
`r unhide()`

9. **Discuss with your neighbour.** The whole design rests on one untestable claim.

   - State the **exclusion restriction** for this instrument in one sentence.
   - Give a concrete reason why the timing of a state's child-labour reform might affect wages
     through something *other* than schooling.
   - Would adding more control variables settle the question?

`r hide("Talking points")`
- **Exclusion restriction:** the required years of schooling in your state and birth year affect
  your later wages *only* by changing how much education you actually got.
- Plausible violations: states reform labour law when their **economy is changing** — industrialising
  states may pass stricter schooling laws *and* offer better-paid jobs later. Reforms may travel
  with other policies (public spending, health, minimum wages). Families may also move between states.
- **No.** Controls close backdoors through variables you measured; the worry here is that the
  *instrument itself* has a second route to wages. More controls cannot rule that out — only an
  argument about how reforms happened can, which is why the paper spends its pages on institutional
  history rather than on regression tables.
`r unhide()`
