---
title: "Exercise 1 (20 minutes)"
subtitle: "You practice: a treatment × country interaction — did one natural experiment hit ten countries the same way? When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 3)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

num_fitb <- function(x, digits = 3, width = 7) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}
```

```{r hidden-read, include = FALSE}
pacman::p_load(tidyverse, haven, estimatr, modelsummary)
ESS_raw <- read_dta("../assets/Legewie_ESS_02.dta")
```

Back to Legewie's Bali natural experiment — but this time across **ten** countries. Was the effect
of the terror attack on xenophobia the *same* everywhere, or did it **depend on the country**? That
is an **interaction** between the treatment and country.

1. Prepare the data. This keeps ten countries and makes **Portugal the reference** (it had the
   clearest effect, so it is the natural baseline to compare the others against).

```{r read, eval = FALSE}
pacman::p_load(tidyverse, haven, estimatr, modelsummary)
ESS_raw <- read_dta("Legewie_ESS_02.dta") # from your project folder
```

```{r prep}
event_date <- as.Date("2002-10-13")   # the Bali attack
win_begin  <- as.Date("2002-09-14")
win_end    <- as.Date("2002-10-20")

ESS <- ESS_raw %>%
  mutate(
    int_date = as.Date(sprintf("%s-%s-%s", inwyr, inwmm, inwdd)),
    treat = case_when(
      int_date > event_date & int_date <= win_end   ~ "After the attack",
      int_date < event_date & int_date > win_begin  ~ "Before the attack",
      TRUE ~ NA_character_) %>% fct_relevel("Before the attack", "After the attack"),
    anti_immi = rowMeans(across(c(imtcjob, imbleco, imbgeco, imueclt,
                                  imwbcnt, imwbcrm, imbghct)), na.rm = TRUE) %>%
      scale() %>% as.numeric(),
    anti_immi = max(anti_immi, na.rm = TRUE) - anti_immi,
    across(c(brncntr, mocntr, facntr), as_factor),
    age  = inwyr - yrbrn,
    empl_stat = case_when(
      pdwrk == 1 ~ "Working", uempla == 1 ~ "Unemployed",
      rtrd == 1 ~ "Retired", TRUE ~ "Other"),
    pspwght = pweight * dweight
  ) %>%
  filter(!cntry %in% c("DK", "IL", "HU") &                # keep 10 countries
           brncntr == "yes" & mocntr == "yes" & facntr == "yes") %>%
  select(treat, anti_immi, cntry, pspwght, age, empl_stat) %>%
  drop_na() %>%
  mutate(cntry = fct_relevel(cntry, "PT"))               # Portugal = reference
```

`r hide("Stuck?")`
- Did `pacman::p_load(...)` run without red errors? Fix that first.
- `read_dta()` can't find the file? `Legewie_ESS_02.dta` must sit **in your project folder**.
- Odd results? *Session → Restart R*, then rerun from the top.
- Still stuck? Ask in the [t-R-ouble forum](https://absalon.ku.dk/courses/70545/discussion_topics/563046?module_item_id=2037268) on Absalon.

`r unhide()`

```{r solutions, include = FALSE}
ols_int <- lm_robust(anti_immi ~ treat * cntry, data = ESS, weights = pspwght)
b <- coef(ols_int)
pt <- b["treatAfter the attack"]
si <- b["treatAfter the attack"] + b["treatAfter the attack:cntrySI"]
```

2. **Predict before you compute.** You are about to fit `anti_immi ~ treat * cntry`. In that model,
   what will the plain `treatAfter the attack` coefficient represent?
   `r mcq(c(answer = "The attack effect in Portugal, the reference country", "The average attack effect across all ten countries", "The attack effect in every country"))`

3. **Fit the interaction** and look at whether the country-specific terms differ from Portugal.

`r hide("Hint")`
`lm_robust(anti_immi ~ treat * cntry, data = ESS, weights = pspwght)` — the `treat:cntryXX` rows are
how much each country's effect differs from Portugal's.
`r unhide()`

`r hide("R solution — don't peek too early ;) !")`
```{r sol3, eval = FALSE}
ols_int <- lm_robust(anti_immi ~ treat * cntry, data = ESS, weights = pspwght)

modelsummary(list("Xenophobia" = ols_int),
             stars = TRUE, gof_map = c("nobs", "r.squared"),
             output = "kableExtra")
```
`r unhide()`

   Compared with Portugal, are the effects in the other countries mostly weaker or stronger?
   `r mcq(c("No significant differences", answer = "Most are significantly smaller", "Most are significantly larger"))`

4. **Read a main term.** What was the effect of the Bali attack **in Portugal** (the reference)?
   `r num_fitb(pt)`

5. **Read a conditional effect.** The effect in **Slovenia** (`SI`) is the Portugal effect *plus*
   the `treatAfter the attack:cntrySI` interaction term. Compute it:
   `r num_fitb(si)`

`r hide("R solution")`
```{r sol5, eval = FALSE}
coef(ols_int)["treatAfter the attack"] +
  coef(ols_int)["treatAfter the attack:cntrySI"]
```
`r unhide()`

6. **Write one sentence** as a `# comment` explaining, in plain words, what a significant
   `treat:cntryXX` interaction *means* — then compare:

`r hide("Model answer")`
A significant `treat:cntryXX` term means the effect of the Bali attack in country XX was
significantly **different** from its effect in Portugal — the attack's impact on xenophobia
**depended on the country**. Because most of these terms are negative and significant, the strong
Portuguese reaction was unusual: in most other countries the same attack moved xenophobia much less,
or not at all. The interaction is what lets one model carry ten different effects at once.
`r unhide()`

7. **Bonus, for the fast.** Add `age` and `empl_stat` as additive controls to the interaction model.
   Do the country differences survive?

`r hide("R solution")`
```{r sol7, eval = FALSE}
ols_int2 <- lm_robust(anti_immi ~ treat * cntry + age + empl_stat,
                      data = ESS, weights = pspwght)

modelsummary(list("No controls" = ols_int, "+ controls" = ols_int2),
             stars = TRUE, gof_map = c("nobs"), output = "kableExtra")
```

The interaction pattern is broadly robust — some countries even become statistically
indistinguishable from Portugal once age and employment are held constant. Note you can freely mix
an **interaction** (`treat * cntry`) with ordinary **additive controls** (`+ age + empl_stat`) in
the same model.
`r unhide()`

8. **Discuss with your neighbour.**

   - Why is it more honest to *interact* treatment with country than to run ten separate
     country-by-country regressions?
   - The Portuguese effect is the reference. If we had made **Sweden** the reference instead, would
     any of the underlying country effects change?
   - A significant treatment × country interaction tells us the effect varies. Does it tell us
     *why* it varies?

`r hide("Talking points")`
- One interaction model uses **all** the data to estimate a common baseline and lets each country
  deviate from it, with proper standard errors for the *differences*. Ten separate regressions throw
  away that shared information and make it awkward to test whether countries really differ.
- **No.** Changing the reference relabels the coefficients — every country's actual predicted effect
  is identical; only which one is called "the baseline" changes. The model is the same.
- **No.** The interaction establishes *that* the effect depends on country, not *why*. Explaining it
  — history, media, prior immigration — is the substantive sociology that begins where the
  regression ends.
`r unhide()`
