---
title: "Exercise 2 (20 minutes)"
subtitle: "You practice: reading a continuous × continuous interaction from BOTH sides — and deciding whether it is real. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 3)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

num_fitb <- function(x, digits = 2, width = 7) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}
```

```{r hidden-setup, include = FALSE}
pacman::p_load(tidyverse, estimatr, vdemdata, wbstats, modelsummary)
```

In the lecture we interacted **state ownership** with **civil liberties** and read the result as
*"the state-ownership slope at different levels of civil liberties"*. But an interaction is
**symmetric** — the very same term also tells us how the **civil-liberties** slope changes with
state ownership. Here you read it from that other side, and then decide whether the pattern is real.

1. Rebuild the Lecture 9 / 11 triangle. (Same pipeline: V-Dem, the World Bank `$3.00` poverty line,
   a cached fallback, and both predictors mean-centred.)

```{r prep}
pacman::p_load(tidyverse, estimatr, vdemdata, wbstats, modelsummary)

Dat_vdem <- vdem %>%
  as_tibble() %>%
  select(country_text_id, year,
         civ_liberties = v2xcl_rol, state_own_raw = v2clstown) %>%
  mutate(state_ownership = -state_own_raw)

Dat_pov <- tryCatch(
  wb_data("SI.POV.DDAY", start_date = 1972, end_date = 2025),
  error = function(e) readRDS("data/wb_poverty_raw.rds")   # offline fallback
) %>%
  rename(poverty = SI.POV.DDAY, year = date, country_text_id = iso3c) %>%
  select(country_text_id, year, country, poverty) %>%
  drop_na(poverty) %>%
  group_by(country) %>% filter(year == max(year)) %>% ungroup()

Dat <- inner_join(Dat_pov, Dat_vdem, by = c("country_text_id", "year")) %>%
  drop_na(poverty, state_ownership, civ_liberties) %>%
  mutate(
    mc_state = state_ownership - mean(state_ownership),
    mc_civ   = civ_liberties  - mean(civ_liberties)
  )
```

`r hide("Stuck?")`
- Did `pacman::p_load(...)` run without red errors? Fix that first.
- The `wb_data()` call needs the internet. If it fails, the `tryCatch` falls back to the cached file
  `data/wb_poverty_raw.rds` — download it from Absalon into a `data/` sub-folder of your project.
- Odd results? *Session → Restart R*, then rerun from the top.
- Still stuck? Ask in the [t-R-ouble forum](https://absalon.ku.dk/courses/70545/discussion_topics/563046?module_item_id=2037268) on Absalon.

`r unhide()`

```{r solutions, include = FALSE}
ols_int <- lm_robust(poverty ~ mc_state * mc_civ, data = Dat)
b <- coef(ols_int)
p_int <- tidy(ols_int) %>% filter(term == "mc_state:mc_civ") %>% pull(p.value)
sd_s <- sd(Dat$mc_state)
civ_lo   <- b["mc_civ"] + b["mc_state:mc_civ"] * (-sd_s)
civ_mean <- b["mc_civ"]
civ_hi   <- b["mc_civ"] + b["mc_state:mc_civ"] * ( sd_s)
```

2. **Fit the interaction** — exactly the lecture model.

`r hide("R solution — don't peek too early ;) !")`
```{r sol2, eval = FALSE}
ols_int <- lm_robust(poverty ~ mc_state * mc_civ, data = Dat)

modelsummary(list("Poverty" = ols_int),
             coef_rename = c("mc_state" = "State ownership",
                             "mc_civ" = "Civil liberties",
                             "mc_state:mc_civ" = "State own. × Civ. lib."),
             stars = TRUE, gof_map = c("nobs", "r.squared"),
             output = "kableExtra")
```
`r unhide()`

   The interaction coefficient is `r num_fitb(b["mc_state:mc_civ"])`. In the lecture we read this as
   "how the **state-ownership** slope changes per unit of civil liberties". Because the interaction
   is symmetric, it *also* means:
   `r mcq(c(answer = "how the civil-liberties slope changes per unit of state ownership", "how much poverty there is on average", "the effect of state ownership when civil liberties are zero"))`

3. **Read the civil-liberties slope at three levels of state ownership.** The civil-liberties slope
   is `b_civ + b_interaction × (state ownership)`. With `mc_state` centred, one standard deviation is
   about `r round(sd_s, 2)`. Compute the slope at low, average and high state ownership:

   - Low state ownership ($-1$ SD): `r num_fitb(civ_lo)`
   - Average state ownership: `r num_fitb(civ_mean)`
   - High state ownership ($+1$ SD): `r num_fitb(civ_hi)`

`r hide("R solution")`
```{r sol3, eval = FALSE}
b    <- coef(ols_int)
sd_s <- sd(Dat$mc_state)

b["mc_civ"] + b["mc_state:mc_civ"] * (-sd_s) # low state ownership
b["mc_civ"]                                  # average
b["mc_civ"] + b["mc_state:mc_civ"] * ( sd_s) # high state ownership
```
`r unhide()`

4. **Draw the predicted lines** — civil liberties on the x-axis, one line per level of state
   ownership. This is the mirror image of the lecture's plot.

`r hide("R solution — don't peek too early ;) !")`
```{r sol4, eval = FALSE}
grid <- expand.grid(
  mc_civ   = seq(min(Dat$mc_civ), max(Dat$mc_civ), length.out = 50),
  mc_state = c(-sd_s, 0, sd_s)
)
grid$poverty <- predict(ols_int, newdata = grid)
grid$Ownership <- factor(grid$mc_state,
  labels = c("Low", "Average", "High"))

ggplot(grid, aes(x = mc_civ, y = poverty, color = Ownership)) +
  geom_line(linewidth = 1.1) +
  labs(x = "Civil liberties (centred)",
       y = "Predicted % below $3.00 a day",
       color = "State ownership") +
  theme_minimal(base_size = 14)
```
`r unhide()`

5. **Now the decision that matters.** Look back at the model table from task 2. Is the interaction
   term statistically significant at the 5% level?
   `r mcq(c("Yes — so the slopes genuinely differ", answer = "No — so we cannot conclude the slopes really differ", "The table does not say"))`

6. **Write one sentence** as a `# comment` stating what you may — and may **not** — conclude from
   this analysis. Then compare:

`r hide("Model answer")`
The fitted lines fan out, which *suggests* that civil liberties are more strongly associated with
lower poverty where state ownership is high. But the interaction term is **not statistically
significant** (p ≈ `r round(p_int, 2)`), so we **cannot** conclude that the civil-liberties slope
genuinely depends on state ownership — the fanning may just be sampling noise. A suggestive picture
is a hypothesis, not a finding. And even had it been significant, this is a cross-country
*association*, not evidence about which variable causes which.
`r unhide()`

7. **Bonus, for the fast.** Confirm the symmetry numerically: the interaction coefficient is
   *identical* whether you write `mc_state * mc_civ` or `mc_civ * mc_state`. Fit both and compare.

`r hide("R solution")`
```{r sol7, eval = FALSE}
coef(lm_robust(poverty ~ mc_state * mc_civ, data = Dat))["mc_state:mc_civ"]
coef(lm_robust(poverty ~ mc_civ * mc_state, data = Dat))["mc_civ:mc_state"]
```

They are the same number: `mc_state:mc_civ` and `mc_civ:mc_state` are one and the same product
column. "Which variable moderates which" is a *story we tell*, not something the maths distinguishes.
`r unhide()`

8. **Discuss with your neighbour.**

   - You read the same interaction two ways (lecture: state-ownership slope by civil liberties;
     here: civil-liberties slope by state ownership). Which framing is "correct"?
   - The interaction was not significant. Would collecting more country-years be a legitimate way to
     "get" a significant result?
   - When *is* a non-significant interaction still worth reporting?

`r hide("Talking points")`
- **Neither is more correct** — they are the same coefficient. Which variable you cast as the
  "moderator" is a rhetorical and theoretical choice, guided by your research question, not by the
  statistics. The maths is perfectly symmetric.
- Collecting more data to shrink standard errors is fine *in principle*, but hunting for
  significance by adding data, dropping cases, or trying specifications until p < 0.05 is
  **p-hacking**. The honest move is to decide the analysis first, then report whatever it gives.
- A non-significant interaction is worth reporting when the question itself matters: "we tested
  whether the effect of state ownership depends on civil liberties and found no reliable evidence
  that it does" is a genuine, publishable answer — and it guards the next researcher against
  over-reading the suggestive picture you just drew.
`r unhide()`
