---
title: "Exercise 2 (20 minutes)"
subtitle: "You practice: the log₂ transformation, comparing it to a polynomial, and visualising the fit. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 3)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

num_fitb <- function(x, digits = 2, width = 7) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}
```

```{r hidden-read, include = FALSE}
pacman::p_load(tidyverse, haven, estimatr, modelsummary)
ESS <- read_spss("../assets/ESS9e03_1.sav") %>%
  filter(cntry == "DK" | cntry == "BG") %>%
  mutate(
    cntry  = as_factor(cntry) %>% fct_relevel("Denmark"),
    gndr   = as_factor(gndr),
    eduyrs = case_when(eduyrs < 9 ~ 9, eduyrs > 21 ~ 21, TRUE ~ zap_labels(eduyrs)),
    agea    = zap_labels(agea),
    imwbcnt = 10 - zap_labels(imwbcnt)
  ) %>%
  select(idno, cntry, pspwght, gndr, eduyrs, agea, imwbcnt) %>%
  drop_na()
```

Same ESS data as exercise 1 (Denmark & Bulgaria; run that setup first). In exercise 1 you fitted a
**polynomial** of age. Now you try the **log₂ transformation** instead — and decide which you prefer.

```{r solutions, include = FALSE}
ols_log  <- lm_robust(imwbcnt ~ cntry + gndr + eduyrs + log2(agea),
                      data = ESS, weights = pspwght)
ols_poly <- lm_robust(imwbcnt ~ cntry + gndr + eduyrs + poly(I(agea / 10), 2, raw = TRUE),
                      data = ESS, weights = pspwght)
b_log <- coef(ols_log)["log2(agea)"]
se_log <- sqrt(diag(vcov(ols_log)))["log2(agea)"]
```

1. **Fit a log₂ model.** Replace the age polynomial with `log2(agea)`, keeping the same controls,
   weighted by `pspwght`.

`r hide("Hint")`
`lm_robust(imwbcnt ~ cntry + gndr + eduyrs + log2(agea), data = ESS, weights = pspwght)`
`r unhide()`

`r hide("R solution — don't peek too early ;) !")`
```{r sol1, eval = FALSE}
ols_log <- lm_robust(imwbcnt ~ cntry + gndr + eduyrs + log2(agea),
                     data = ESS, weights = pspwght)

modelsummary(list("Xenophobia" = ols_log),
             coef_rename = c("log2(agea)" = "Age (log2)"),
             stars = TRUE, gof_map = c("nobs", "r.squared"), output = "kableExtra")
```
`r unhide()`

2. The `log2(agea)` coefficient is `r num_fitb(b_log)`. Because the predictor is on a **log₂**
   scale, this reads as: a **doubling of age** (say from 20 to 40, or 40 to 80) is associated with
   about `r num_fitb(b_log)` scale points
   `r mcq(c(answer = "more xenophobia", "less xenophobia", "no change in xenophobia"))`.

3. **Compare the two curves.** Put the log₂ model next to the quadratic from exercise 1 and look at
   the fit statistics.

`r hide("R solution — don't peek too early ;) !")`
```{r sol3, eval = FALSE}
ols_poly <- lm_robust(imwbcnt ~ cntry + gndr + eduyrs + poly(I(agea / 10), 2, raw = TRUE),
                      data = ESS, weights = pspwght)

modelsummary(list("Quadratic" = ols_poly, "log2" = ols_log),
             gof_map = c("nobs", "r.squared", "adj.r.squared"),
             stars = TRUE, output = "kableExtra")
```
`r unhide()`

   According to $R^2$ and adjusted $R^2$, which model fits better?
   `r mcq(c("The quadratic is clearly better", "The log₂ is clearly better", answer = "They are essentially equally good"))`

4. Given that the two fit about equally well, which would you **report** — and why?
   `r mcq(c(answer = "The log₂ model — it is simpler (one coefficient) and reads as 'per doubling of age'", "The quadratic — more coefficients always means a better model", "It does not matter at all which you report"))`

5. **Write one sentence** as a `# comment` justifying your choice — then compare:

`r hide("Model answer")`
The log₂ and quadratic models fit the age–xenophobia relationship about equally well (nearly
identical $R^2$ and adjusted $R^2$). When fit is a tie, prefer the **simpler, more interpretable**
model: the log₂ needs a single coefficient and reads cleanly as "a doubling of age is associated
with about 0.4 points more xenophobia", whereas the quadratic needs two coefficients and a turning
point to interpret. The one reason to prefer the quadratic would be a genuine **rise-then-fall** in
age — which we did not find here.
`r unhide()`

6. **Visualise the log₂ fit.** Make predictions across the age range and plot the implied curve.

`r hide("R solution")`
```{r sol6, eval = FALSE}
# 1. Fictional data spanning the observed ages, controls held constant
fict <- tibble(
  agea   = seq(15, 90, by = 1),
  cntry  = "Denmark",
  gndr   = "Female",
  eduyrs = mean(ESS$eduyrs)
)

# 2. Predicted xenophobia + 95% interval
fict <- predict(ols_log, newdata = fict, interval = "confidence")$fit %>%
  as_tibble() %>%
  bind_cols(fict, .)

# 3. Plot
ggplot(fict, aes(x = agea, y = fit)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.25) +
  geom_line(linewidth = 1, colour = "#901A1E") +
  labs(x = "Age", y = "Predicted xenophobia (0–10)") +
  theme_minimal()
```

The curve rises quickly among the young and flattens among the old — the signature shape of a log
predictor.
`r unhide()`

7. **Bonus, for the fast.** The log of a variable is only defined for **positive** values. Which
   variables in a typical survey could you *not* log without a fix — and what is the usual trick?

`r hide("Talking points")`
`log()` is undefined at 0 and for negatives. Income with zeros, counts that include 0, or anything
centred around 0 cannot be logged directly. Common fixes: add a small constant (`log(x + 1)`, i.e.
`log1p`), or use a different transformation. Age is safe because everyone is older than zero.
`r unhide()`

8. **Discuss with your neighbour.**

   - The deck's health-spending example was a case where **log₂ clearly won**. Here it only **ties**
     the polynomial. What is different about the two relationships?
   - "The coefficient on `log2(agea)` is 0.4." Why is *"per doubling"* a more honest way to say this
     than *"per year"*?
   - If you logged the **outcome** instead of the predictor, how would the interpretation change?

`r hide("Talking points")`
- Health spending spans a huge range (from a few dollars to over \$10,000) and flattens sharply, so
  the log dramatically straightens it. Age spans a much narrower range (15–90, barely more than two
  doublings) and bends only gently, so both tools capture the mild curve about equally.
- A log predictor means the effect is **not** constant per year — it is larger among the young and
  smaller among the old. "Per doubling" states that honestly; "per year" would falsely imply a
  constant effect, which is exactly the linearity assumption the log was meant to relax.
- Logging the *outcome* makes coefficients read as **approximate percentage** changes in $Y$ per
  unit of $X$ — the convention you already met with log wages in Lecture 10.
`r unhide()`
