---
title: "Exercise 2 (20 minutes)"
subtitle: "Edge ③ — and the verdict on the trade-off. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

pacman::p_load(
  tidyverse,    # Data manipulation & figures
  wbstats,      # World Bank data via API
  vdemdata,     # Varieties of Democracy data
  estimatr,     # OLS with robust standard errors
  modelsummary  # Nicely formatted regression tables
)

# 1. V-Dem: freedom, and state control of the economy
Dat_vdem <- vdem %>%
  as_tibble() %>%
  select(country_text_id, year,
         equal_liberty = v2xcl_rol,
         state_owner   = v2clstown) %>%
  mutate(state_control = -state_owner)

# 2. World Bank: extreme poverty (share below $3.00 a day, 2021 PPP)
Dat_poverty <- wb_data("SI.POV.DDAY", start_date = 1972, end_date = 2025) %>%
  rename(poverty = SI.POV.DDAY,
         year = date,
         country_text_id = iso3c) %>%
  select(country_text_id, year, country, poverty) %>%
  drop_na(poverty) %>%
  group_by(country) %>%
  filter(year == max(year)) %>%
  ungroup()

# 3. Join
Dat <- inner_join(Dat_poverty, Dat_vdem, by = c("country_text_id", "year"))
```

```{r echo = FALSE}
ols_1 <- lm_robust(poverty ~ state_control, data = Dat)
ols_2 <- lm_robust(poverty ~ equal_liberty, data = Dat)
```

*(Keep working in the R script from Exercise 1 — you already have `Dat`.)*

1. Regress poverty on state control of the economy (edge ① again, now as a regression).
   With a one-unit increase in state control, poverty changes by:
   `r fitb(round(coef(ols_1)[2], 2), width = "6")` percentage points (two digits).

`r hide("R solution -> don't peek too early ;) !")`
```{r results = 'asis'}
ols_1 <- lm_robust(poverty ~ state_control, data = Dat)

modelsummary(
  list("Poverty" = ols_1),
  statistic = NULL,                 # No statistical inference (yet)
  gof_map = c("nobs", "r.squared")
)
```
`r unhide()`

2. Regress poverty on the index of equality and liberty (edge ② as a regression).
   With a one-unit increase in equality and liberty rights, poverty changes by:
   `r fitb(round(coef(ols_2)[2], 2), width = "6")` percentage points (two digits).

`r hide("R solution -> don't peek too early ;) !")`
```{r results = 'asis'}
ols_2 <- lm_robust(poverty ~ equal_liberty, data = Dat)

modelsummary(
  list("Poverty" = ols_2),
  statistic = NULL,
  gof_map = c("nobs", "r.squared")
)
```
`r unhide()`

3. Which of the two is the better model of the variance of poverty across the world?
`r mcq(c("The one with state control of the economy", answer = "The one with equality and liberty rights, as the R2 shows", "The one with equality and liberty rights, as the stronger beta coefficient shows"))`

4. Now the missing edge ③: make a scatter plot with the index of equality and liberty
   on the Y-axis and state control of the economy on the X-axis.

`r hide("R solution -> don't peek too early ;) !")`
```{r out.width='100%', fig.height = 4, fig.width = 7.5, results = FALSE, echo = TRUE}
ggplot(data = Dat, aes(y = equal_liberty, x = state_control)) +
  geom_text(aes(label = country)) +
  labs(y = "Index of equality before the law and individual liberty",
       x = "State control of the economy (V-Dem, reversed)") +
  theme_minimal()
```
`r unhide()`

```{r echo = FALSE}
ols_3 <- lm_robust(equal_liberty ~ state_control, data = Dat)
```

5. Regress the index of equality and liberty on state control. With a one-unit increase
   in state control, equality and liberty rights change by:
   `r fitb(round(coef(ols_3)[2], 2), width = "6")` (two digits).

`r hide("R solution -> don't peek too early ;) !")`
```{r results = 'asis'}
ols_3 <- lm_robust(equal_liberty ~ state_control, data = Dat)

modelsummary(
  list("Equality & liberty" = ols_3),
  statistic = NULL,
  gof_map = c("nobs", "r.squared")
)
```
`r unhide()`

6. Make one regression table with all three models — the complete triangle. Based on the
   table, **discuss with your neighbour:** is there a freedom–equality trade-off? Which
   edge of the triangle is strong, which is weak, and what does that mean?

`r hide("R solution -> don't peek too early ;) !")`
```{r results = 'asis'}
modelsummary(
  list("Poverty" = ols_1, "Poverty" = ols_2, "Equality & liberty" = ols_3),
  statistic = NULL,
  gof_map = c("nobs", "r.squared")
)
```
`r unhide()`
