---
title: "Exercise 1 (20 minutes)"
subtitle: "Edge ② of our triangle: freedom and poverty. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
```

1. Do the [Setup](about.html), if you have not yet done so.

2. Start a new R script and copy the lines of R code from the slides that generate `Dat`:
   get the V-Dem data, download the World Bank poverty data via the API, and join the two.

`r hide("R solution -> don't peek too early ;) !")`
```{r}
pacman::p_load(
  tidyverse,    # Data manipulation & figures
  wbstats,      # World Bank data via API
  vdemdata,     # Varieties of Democracy data
  estimatr,     # OLS with robust standard errors
  modelsummary  # Nicely formatted regression tables
)

# 1. V-Dem: freedom, and state control of the economy
Dat_vdem <- vdem %>%
  as_tibble() %>%
  select(country_text_id, year,
         equal_liberty = v2xcl_rol,    # Equality before the law & individual liberty
         state_owner   = v2clstown) %>% # State ownership of the economy
  # v2clstown runs from "all state-owned" (low) to "all private" (high),
  # so we flip it: higher = MORE state control
  mutate(state_control = -state_owner)

# 2. World Bank: extreme poverty (share below $3.00 a day, 2021 PPP)
Dat_poverty <- wb_data("SI.POV.DDAY", start_date = 1972, end_date = 2025) %>%
  rename(poverty = SI.POV.DDAY,
         year = date,
         country_text_id = iso3c) %>%
  select(country_text_id, year, country, poverty) %>%
  drop_na(poverty) %>%
  group_by(country) %>%
  filter(year == max(year)) %>% # Keep the most recent estimate per country
  ungroup()

# 3. Join the two data sets by country and year
Dat <- inner_join(Dat_poverty, Dat_vdem, by = c("country_text_id", "year"))
```
`r unhide()`

3. Make a scatter plot with poverty as outcome on the Y-axis and the index of equality
   and liberty as predictor on the X-axis. Remember the four questions to ask a scatter plot:
   direction, form, spread, outliers.

`r hide("R solution -> don't peek too early ;) !")`
```{r out.width='100%', fig.height = 4, fig.width = 7.5, results = FALSE, echo = TRUE}
ggplot(data = Dat, aes(y = poverty, x = equal_liberty)) +
  geom_text(aes(label = country)) +
  labs(y = "% of population below $3.00 a day (2021 PPP)",
       x = "Index of equality before the law and individual liberty") +
  theme_minimal()
```
`r unhide()`

```{r results = FALSE, echo = FALSE}
cor_povlib <- Dat %>%
  select(poverty, equal_liberty) %>%
  cor() %>%
  round(2)
cor_povlib <- cor_povlib[1, 2]
```

4. Estimate the correlation between poverty and the index of equality and liberty
   (rounded to two digits): `r fitb(cor_povlib, width = "5")`

`r hide("R solution -> don't peek too early ;) !")`
```{r}
Dat %>%
  select(poverty, equal_liberty) %>%
  cor() %>%
  round(2)
```
`r unhide()`

5. This is edge ② of our triangle. What does the correlation suggest about a
   freedom–equality trade-off?
`r mcq(c(answer = "No trade-off visible here: where people have more equality and liberty rights, there tends to be less poverty", "A trade-off: freedom comes at the price of more poverty", "The result says nothing about the relationship"))`
