---
title: "Exercise 1 (20 minutes)"
subtitle: "You practice: APIs, joins, scatter plots, and the correlation coefficient — edge ② of our triangle. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

# Fill-in-the-blank that accepts point AND comma decimals (Danish keyboards!)
num_fitb <- function(x, digits = 2, width = 6) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}
```

1. Do the [Setup](about.html), if you have not yet done so.

2. Start a new R script and copy in this starter code (it loads the packages and
   downloads both data sets — we walked through every line on the slides):

```{r}
pacman::p_load(
  tidyverse,    # Data manipulation & figures
  wbstats,      # World Bank data via API
  vdemdata,     # Varieties of Democracy data
  estimatr,     # OLS with robust standard errors
  modelsummary  # Nicely formatted regression tables
)

# V-Dem: civil liberties, and state ownership of the economy
Dat_vdem <- vdem %>%
  as_tibble() %>%
  select(country_text_id, year,
         civ_liberties = v2xcl_rol,    # Equality before the law & individual liberty
         state_own_raw   = v2clstown) %>% # State ownership of the economy
  mutate(state_ownership = -state_own_raw)  # Reverse: higher = MORE state ownership

# World Bank: extreme poverty (share below $3.00 a day, 2021 PPP).
# Try the live API; if it is down, fall back to the cached copy.
Dat_poverty <- tryCatch(
  wb_data("SI.POV.DDAY", start_date = 1972, end_date = 2025),
  error = function(e) readRDS("data/wb_poverty_raw.rds")
) %>%
  rename(poverty = SI.POV.DDAY,
         year = date,
         country_text_id = iso3c) %>%
  select(country_text_id, year, country, poverty) %>%
  drop_na(poverty) %>%
  group_by(country) %>%
  filter(year == max(year)) %>% # Keep the most recent estimate per country
  ungroup()
```

3. **Now write the join yourself:** combine `Dat_poverty` and `Dat_vdem` into one tibble
   called `Dat`. Think first: *which* join do you want, and *which variables form the key*?

`r hide("Hint")`
We want only countries that appear in **both** tibbles (an *inner* join), and a country
is identified by its ISO code *together with* the year.
`r unhide()`

`r hide("R solution -> don't peek too early ;) !")`
```{r}
Dat <- inner_join(Dat_poverty, Dat_vdem, by = c("country_text_id", "year"))
```
`r unhide()`

> **Stuck?**
>
> - Did the `pacman::p_load(...)` line run without red error messages? If not, fix that first.
> - Strange results? Clear leftovers with a fresh start: *Session → Restart R*, then rerun your script from the top.
> - **World Bank API down or very slow?** Load our cached copy instead:
>   `Dat <- readRDS(url("https://willowy-quokka-e604ee.netlify.app/data/Dat_L2.rds"))`
> - Still stuck? Ask (and answer!) in the [t-R-ouble forum](https://absalon.ku.dk/courses/70545/discussion_topics/563046?module_item_id=2037268) on Absalon.

4. Make a scatter plot with poverty as outcome on the Y-axis and civil liberties as
   predictor on the X-axis. Ask it the four questions from the lecture:
   *direction, form, spread, outliers*.

`r hide("R solution -> don't peek too early ;) !")`
```{r out.width='100%', fig.height = 4, fig.width = 7.5, results = FALSE, echo = TRUE}
ggplot(data = Dat, aes(y = poverty, x = civ_liberties)) +
  geom_text(aes(label = country)) +
  labs(y = "% of population below $3.00 a day (2021 PPP)",
       x = "Civil liberties") +
  theme_minimal()
```
`r unhide()`

5. **Before you compute anything:** judging only from your scatter plot, the correlation
   between poverty and civil liberties is …
`r mcq(c("close to −0.8", answer = "close to −0.3", "close to 0", "close to +0.3"))`

```{r results = FALSE, echo = FALSE}
cor_povlib <- Dat %>%
  select(poverty, civ_liberties) %>%
  cor()
cor_povlib <- cor_povlib[1, 2]
```

6. Now check yourself: estimate the correlation with `cor()` (rounded to two digits;
   point or comma is fine): `r num_fitb(cor_povlib)`

`r hide("R solution -> don't peek too early ;) !")`
```{r}
Dat %>%
  select(poverty, civ_liberties) %>%
  cor() %>%
  round(2)
```
`r unhide()`

7. This is edge ② of our triangle. What does the correlation suggest about a
   liberty–equality trade-off?
`r mcq(c(answer = "No trade-off visible here: where people have more civil liberties, there tends to be less poverty", "A trade-off: stronger civil liberties come at the price of more poverty", "The result says nothing about the relationship"))`

8. **Write one sentence** in your R script (as a `# comment`) interpreting the correlation —
   as if for your term paper. Then compare with the model answer:

`r hide("Model answer")`
*Across `r nrow(Dat)` countries, civil liberties correlate moderately negatively with extreme poverty (r ≈ `r round(cor_povlib, 2)`): countries with stronger civil liberties tend to have a smaller share of their population in extreme poverty.*

(Yours doesn't need the same words — but it should name the **direction**, the rough
**strength**, and the **units of analysis** (countries), and it should *not* claim causality.)
`r unhide()`

------------------------------------------------------------------------------

**Bonus (for the fast):** `Dat` has ~160 countries, but V-Dem covers 179 — where did the
rest go? Find the countries we *lost* in the join, using `anti_join()`. Why are they missing?

`r hide("R solution")`
```{r}
Dat_vdem %>%
  filter(year > 2015) %>% # Recent years only
  anti_join(Dat_poverty, by = c("country_text_id", "year")) %>%
  distinct(country_text_id)
```

They are countries (and country-years) with **no World Bank poverty estimate** — poverty
data require household surveys, which some countries run rarely or not at all (e.g. North
Korea). An `inner_join()` silently drops them. **Data that are missing are a finding, too:**
our analysis can only speak about countries that measure poverty.
`r unhide()`
