---
title: "Exercise 2 (20 minutes)"
subtitle: "You practice: OLS regression and its interpretation — edge ③, and the verdict on the trade-off. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

# Fill-in-the-blank that accepts point AND comma decimals (Danish keyboards!)
num_fitb <- function(x, digits = 2, width = 6) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}

pacman::p_load(
  tidyverse,    # Data manipulation & figures
  wbstats,      # World Bank data via API
  vdemdata,     # Varieties of Democracy data
  estimatr,     # OLS with robust standard errors
  modelsummary  # Nicely formatted regression tables
)

# V-Dem: civil liberties, and state ownership of the economy
Dat_vdem <- vdem %>%
  as_tibble() %>%
  select(country_text_id, year,
         civ_liberties = v2xcl_rol,
         state_own_raw   = v2clstown) %>%
  mutate(state_ownership = -state_own_raw)

# World Bank: extreme poverty (share below $3.00 a day, 2021 PPP).
# Try the live API; if it is down, fall back to the cached copy.
Dat_poverty <- tryCatch(
  wb_data("SI.POV.DDAY", start_date = 1972, end_date = 2025),
  error = function(e) readRDS("data/wb_poverty_raw.rds")
) %>%
  rename(poverty = SI.POV.DDAY,
         year = date,
         country_text_id = iso3c) %>%
  select(country_text_id, year, country, poverty) %>%
  drop_na(poverty) %>%
  group_by(country) %>%
  filter(year == max(year)) %>%
  ungroup()

# Join
Dat <- inner_join(Dat_poverty, Dat_vdem, by = c("country_text_id", "year"))
```

```{r echo = FALSE}
ols_1 <- lm_robust(poverty ~ state_ownership, data = Dat)
ols_2 <- lm_robust(poverty ~ civ_liberties, data = Dat)
```

*(Keep working in the R script from Exercise 1 — you already have `Dat`. If not: the
starter code and the "Stuck?" box are in Exercise 1.)*

1. Regress poverty on state ownership of the economy (edge ① again, now as a regression).
   With a one-unit increase in state ownership, poverty changes by:
   `r num_fitb(coef(ols_1)[2])` percentage points (two digits; point or comma is fine).

`r hide("R solution -> don't peek too early ;) !")`
```{r results = 'asis'}
ols_1 <- lm_robust(poverty ~ state_ownership, data = Dat)

modelsummary(
  list("Poverty" = ols_1),
  statistic = NULL,                 # No statistical inference (yet)
  gof_map = c("nobs", "r.squared")
)
```
`r unhide()`

2. **Write one sentence** in your R script (as a `# comment`) interpreting that slope —
   *descriptively*, as in the lecture. Then compare:

`r hide("Model answer")`
*Across `r nrow(Dat)` countries, one unit more state ownership of the economy (≈ one
standard deviation on the V-Dem scale) is associated with, on average,
`r round(coef(ols_1)[2], 2)` percentage points more of the population in extreme poverty.*

(Direction, size, units — and no causal language: "is associated with", not "leads to".)
`r unhide()`

3. Regress poverty on civil liberties (edge ② as a regression).
   With a one-unit increase in civil liberties, poverty changes by:
   `r num_fitb(coef(ols_2)[2])` percentage points (two digits).

`r hide("R solution -> don't peek too early ;) !")`
```{r results = 'asis'}
ols_2 <- lm_robust(poverty ~ civ_liberties, data = Dat)

modelsummary(
  list("Poverty" = ols_2),
  statistic = NULL,
  gof_map = c("nobs", "r.squared")
)
```

*One sentence: countries with a one-unit stronger civil liberties have, on
average, `r abs(round(coef(ols_2)[2], 1))` percentage points less of their population in
extreme poverty.*
`r unhide()`

4. Which of the two is the better model of the variance of poverty across the world?
`r mcq(c("The one with state ownership of the economy", answer = "The one with civil liberties, as the R2 shows", "The one with civil liberties, as the stronger beta coefficient shows"))`

5. Now the missing edge ③: make a scatter plot with civil liberties
   on the Y-axis and state ownership of the economy on the X-axis.

`r hide("R solution -> don't peek too early ;) !")`
```{r out.width='100%', fig.height = 4, fig.width = 7.5, results = FALSE, echo = TRUE}
ggplot(data = Dat, aes(y = civ_liberties, x = state_ownership)) +
  geom_text(aes(label = country)) +
  labs(y = "Civil liberties",
       x = "State ownership of the economy (V-Dem, reversed)") +
  theme_minimal()
```
`r unhide()`

6. **Before computing:** judging from your scatter plot, the correlation on edge ③ is …
`r mcq(c(answer = "close to −0.7", "close to −0.2", "close to 0", "close to +0.4"))`

```{r echo = FALSE}
ols_3 <- lm_robust(civ_liberties ~ state_ownership, data = Dat)
```

7. Regress civil liberties on state ownership. With a one-unit increase
   in state ownership, civil liberties change by:
   `r num_fitb(coef(ols_3)[2])` (two digits).

`r hide("R solution -> don't peek too early ;) !")`
```{r results = 'asis'}
ols_3 <- lm_robust(civ_liberties ~ state_ownership, data = Dat)

modelsummary(
  list("Civil liberties" = ols_3),
  statistic = NULL,
  gof_map = c("nobs", "r.squared")
)
```
`r unhide()`

8. Make one regression table with all three models — the complete triangle — and
   **discuss with your neighbour**, in three steps:

   a. Which edge of the triangle shows the *strongest* association, which the *weakest*?
   b. A defender of the trade-off thesis points at China (state-owned, extreme
      poverty ≈ 0). What do you answer, looking at *all* countries?
   c. Why can none of this be read *causally* (yet)? What would we need?

`r hide("R solution + talking points")`
```{r results = 'asis'}
modelsummary(
  list("Poverty" = ols_1, "Poverty" = ols_2, "Civil liberties" = ols_3),
  statistic = NULL,
  gof_map = c("nobs", "r.squared")
)
```

*Talking points:* **(a)** Edge ③ is by far the strongest — state ownership comes with far
fewer civil liberties; edge ① is the weakest — no systematic poverty payoff. **(b)** One case is an
anecdote; across all countries the average pattern shows no poverty advantage. (And is
China's poverty reduction *because of* state ownership, or of markets, growth, or something
else entirely?) **(c)** These are descriptive comparisons of *different* countries that
differ in many further ways — confounding. What we would need is the topic of the coming
weeks: designs and methods that identify causal effects.
`r unhide()`

------------------------------------------------------------------------------

**Bonus (for the fast):** rerun edge ① *without* China and Vietnam. Does the conclusion
change? What does that tell you about how sensitive our verdict is to single cases?

`r hide("R solution")`
```{r results = 'asis'}
ols_1b <- lm_robust(poverty ~ state_ownership,
                    data = Dat %>% filter(!country %in% c("China", "Viet Nam")))

modelsummary(
  list("All countries" = ols_1, "Without CHN & VNM" = ols_1b),
  statistic = NULL,
  gof_map = c("nobs", "r.squared")
)
```

The slope barely moves — the "no poverty payoff" conclusion does **not** hinge on the two
famous cases. Checking whether results survive removing influential cases is called a
*robustness check* — remember it for your term paper.
`r unhide()`
