---
title: "Exercise 2 (20 minutes)"
subtitle: "You practice: standard errors, t- and p-values, and confidence intervals. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

# Fill-in-the-blank that accepts point AND comma decimals (Danish keyboards!)
num_fitb <- function(x, digits = 3, width = 6) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}

library(tidyverse); library(haven); library(estimatr); library(modelsummary)
ESS <- read_spss("../assets/ESS9e03_1.sav") %>%
  filter(cntry == "DK") %>%
  select(idno, pspwght, gndr, eduyrs, agea, psppsgva, trstlgl, trstplc) %>%
  mutate(across(c(psppsgva, trstlgl, trstplc, pspwght), zap_labels),
         eduyrs = pmin(pmax(eduyrs, 9), 21),
         gndr = as_factor(gndr)) %>%
  drop_na()

mod  <- lm_robust(trstlgl ~ eduyrs, data = ESS, weights = pspwght)
res  <- tidy(mod) %>% filter(term == "eduyrs")
b    <- res$estimate; se <- res$std.error
tval <- res$statistic; pval <- res$p.value
```

Now the same logic, on **trust in the legal system** (`trstlgl`). Keep the prepared `ESS`
from Exercise 1 (its "Stuck?" box still applies).

1. **Write it yourself:** regress trust in the legal system on years of education, **with
   weights**, and print the full output with `summary()`.

`r hide("R solution -> don't peek too early ;) !")`
```{r sol1}
mod <- lm_robust(trstlgl ~ eduyrs, data = ESS, weights = pspwght)
summary(mod)
```
`r unhide()`

2. Read off the estimated slope $\hat\beta$: one more year of education goes with a change
   in legal-system trust of `r num_fitb(b)` (three digits; point or comma is fine).

3. What is its estimated **standard error** $\hat\sigma$? `r num_fitb(se)`

4. The **$t$-value** here is about `r sprintf("%.1f", tval)`. What does it tell you?
`r mcq(c("The estimate is too large to be significant", "The probability that H0 is true is about 5%", answer = "The estimate is that many standard errors away from 0", "The share of variance the model explains"))`

5. And the **$p$-value**? Which reading is correct?
`r mcq(c("The probability that H0 is true", "The probability that education really predicts trust", answer = "If H0 (β = 0) were true, the probability of drawing a sample with an estimate this far from 0 (or further)", "The share of respondents who trust the legal system"))`

6. **Predict, then check.** With $p =$ `r format.pval(pval, digits = 2)`, at the conventional
   5% threshold we …
`r mcq(c(answer = "reject H0 — the association is statistically significant", "cannot reject H0 — the association could be chance", "have proven that education causes trust"))`

7. **Write two sentences** (as `# comments`): (a) can we reject $H_0$, and (b) does that
   make the effect *causal*? Then compare:

`r hide("Model answer")`
*(a) Yes: p is well below 0.05, so we reject the null of no association — better-educated
Danes trust the legal system significantly more. (b) No — this is a significant
**association** in observational survey data. Significance rules out chance, not
confounding: more-educated people differ in many other ways too. Causal claims need the
designs we meet later in the course.*
`r unhide()`

8. Visualise it: make a scatter plot of `trstlgl` on `eduyrs` with a weighted OLS line and
   its **95% confidence band**, and map the weights to dot size.

`r hide("R solution -> don't peek too early ;) !")`
```{r sol2, out.width='70%', fig.height = 4, fig.width = 6.5, results = FALSE}
ggplot(ESS, aes(y = trstlgl, x = eduyrs)) +
  geom_jitter(aes(size = pspwght), alpha = 1/3, height = 0.1, width = 0.1) +
  geom_smooth(aes(weight = pspwght), method = "lm", color = "#901A1E") +
  labs(y = "Trust in the legal system", x = "Years of education") +
  theme_minimal() +
  theme(legend.position = "none")
```
`r unhide()`

------------------------------------------------------------------------------

**Bonus (for the fast):** re-estimate the model on a **random sample of 50** respondents
(`ESS %>% sample_n(50)`). Is the slope still significant? Run it a few times (without
`set.seed`) — what does that teach you about sample size and sampling error?

`r hide("R solution + takeaway")`
```{r bonus}
mod_small <- lm_robust(trstlgl ~ eduyrs, data = ESS %>% sample_n(50))
tidy(mod_small) %>% filter(term == "eduyrs") %>%
  select(estimate, std.error, statistic, p.value)
```

With only 50 people the standard error is large, so the estimate usually is **not**
significant — and it jumps around from run to run. The full-sample effect is real, but a
tiny sample rarely has the *power* to detect it. **More data → smaller standard error →
sharper conclusions.**
`r unhide()`
