---
title: "Exercise 2 (20 minutes)"
subtitle: "You practice: DAGs and a balance test — is racial appearance balanced across news readers? When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

pacman::p_load(tidyverse, estimatr, modelsummary)

# Load and prepare APAD (see Exercise 1's "Stuck?" box if load() fails)
load("../assets/APAD.RData")
APAD <- APAD %>%
  mutate(
    news    = news_hrs * 60 + news_mins,
    news_yn = if_else(news >= 15, 1, 0)
  )
```

*(Keep working in the R script from Exercise 1 — you already have `APAD` prepared.)*

1. **Draw a DAG with a neighbour** (pen and paper is fine). Start with two nodes — **News
   consumption** and **Support for an anti-discrimination law** — and an arrow from the first to
   the second. Now skim the APAD variables and add any **confounders** you can think of: a
   confounder needs an arrow into *both* nodes.

`r hide("Some candidates")`
Age, German citizenship, immigrant generation, education, or racial appearance could all plausibly
shape *both* how much news someone follows *and* their attitudes toward anti-discrimination law —
each one opens a backdoor path.
`r unhide()`

2. **Write it yourself: a balance test.** The variable `appearance` records how a respondent's
   looks are typically described in Germany (White, Black, Asian, Arabic, Southern). Test whether
   `appearance` is balanced across news readers vs. non-readers.

`r hide("Hint")`
`select()` just `news_yn`, `appearance`, and `gewFAKT`; `rename(weights = gewFAKT)` so it is treated as a weight; then `datasummary_balance(~ news_yn, data = .)`.
`r unhide()`

`r hide("R solution -> don't peek too early ;) !")`
```{r results = 'asis'}
APAD %>%
  select(news_yn, appearance, gewFAKT) %>%
  rename(weights = gewFAKT) %>%
  datasummary_balance(~ news_yn, data = .,
                      title = "Racial appearance of news readers vs. non-readers")
```
`r unhide()`

3. Read your balance table. What does it tell you?
`r mcq(c("Appearance is perfectly balanced, so it cannot bias the comparison", answer = "The groups differ in their racial-appearance mix — a sign the comparison is confounded", "The table proves that appearance causes news reading"))`

4. **Write one sentence** (as a `# comment`) explaining *why* an imbalance in `appearance` matters
   for our estimate. Then compare:

`r hide("Model answer")`
*If people who are racialised differently both follow the news differently **and** face different
levels of discrimination (and hold different views on anti-discrimination law), then appearance
sits on a backdoor path — the raw news–attitude difference is partly selection bias, not a causal
effect.*
`r unhide()`

------------------------------------------------------------------------------

5. **Discuss with your neighbour** — the limits of a balance test:

   a. A balance test can only check variables we **measured**. Name two **unobserved** confounders
      that are *not* in the APAD data but could bias the news effect.
   b. For each, say how it might shape *news consumption* and *support for the law*.
   c. Why can no balance test ever *prove* that a comparison is unbiased?

`r hide("Talking points")`
**(a/b)** For example: political interest (drives more news *and* stronger pro-law attitudes);
experiences of discrimination from family (shape both media diet and attitudes); trust in
institutions; personality. **(c)** Balance on *observed* variables says nothing about the
*unobserved* ones — and the baseline $Y_0$ we truly care about is never seen. Only a design that
makes the groups comparable *by construction* (next week: **randomized experiments**) can close
every backdoor at once.
`r unhide()`
