---
title: "Exercise 2 (20 minutes)"
subtitle: "You practice: estimating causal effects from an RCT, reading standard errors, and coefficient plots. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(tidyverse)

# Fill-in-the-blank that accepts point AND comma decimals (Danish keyboards!)
num_fitb <- function(x, digits = 3, width = 6) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}

# 1. Load packages
pacman::p_load( # Load several R packages using the pacman package manager
  tidyverse,  # A collection of packages for data manipulation and visualization
  ggplot2,    # Powerful package for creating static, animated and interactive visualizations
  estimatr,   # Package for fast estimators for regression with weighted data
  modelr,     # Provides functions for modelling and prediction
  kableExtra, # Enhances table creation in R
  modelsummary) # Creates tables and plots to summarize statistical models

# 2. read the APAD data
load("../assets/APAD.RData") # Read APAD data,

# 3. Process APAD data
APAD <- APAD %>% mutate( # Process APAD data
  # Convert news consumption to minutes
  # Example: 2 hours and 30 minutes becomes 2*60 + 30 = 150 minutes
  news = news_hrs * 60 + news_mins,
  # Create binary variable for news consumption
  news_yn = case_when(
    news < 15 ~ 0,  # 0 if less than 15 minutes
    news >= 15 ~ 1, # 1 if 15 minutes or more
    TRUE ~ as.numeric(NA)), # Handle other cases as missing
  # Calculate average discrimination index across multiple domains
  dis_index = rowMeans( # Calculate the mean for each row (participant)
    select(., # Choose specific discrimination columns from "." (APAD)
           dis_trainee, dis_job, dis_school,
           dis_house, dis_gov, dis_public),
    na.rm = TRUE), # Ignore NA values
  # Standardize discrimination index
  # scale() standardizes; as.numeric() converts to numeric vector
  z_dis_index = scale(dis_index) %>% as.numeric()
)
```

*(Keep working in the R script from Exercise 1 — you already have the prepared `APAD` data.
The "Stuck?" box is in Exercise 1.)*

1. The APAD data contains a variable `antidiscr_law`. It asks respondents how good or bad they think it is for a country to have a law against ethnic discrimination in the workplace. The answer categories are: (1) Very bad, (2) Bad, (3) So/so, (4) Good, (5) Very good. <br> Use the APAD survey experiment to estimate the causal effect of reading a news article about discrimination on support for anti-discrimination policy. Estimate z-standardized effects.
`r hide("R solution -> don't peek too early ;) !")`
```{r}
APAD <- APAD %>%
  mutate( # Standardize 'antidiscr_law' variable
    z_antidiscr_law = scale(antidiscr_law) %>% as.numeric())

zols <- lm_robust( # Weighted OLS on the survey experiment
  z_antidiscr_law ~ article,
  weights = gewFAKT,
  data = APAD
)

modelsummary( # Regression table with stars and fit statistics
  list("Anti-discr. law" = zols),
  stars = TRUE,
  gof_map = c("nobs", "r.squared")
)
```
`r unhide()`

```{r include = FALSE}
coef_1 <- coef(zols)["articleTreat_1"] %>% round(3) %>% unname()
se_1 <- vcov(zols) %>% diag()
se_1 <- se_1["articleTreat_1"] %>% sqrt() %>% round(3) %>% unname()
```

2. What is the causal effect of reading news about discrimination on supporting anti-discrimination policy? The causal effect is: `r num_fitb(coef_1)` (three digits; point or comma is fine). What is the estimated standard error of this $\hat{\beta}$? `r num_fitb(se_1)`. *What do these two statistics suggest about a causal effect?*

`r hide("Model answer")`
*Reading the discrimination article increased support for anti-discrimination law by
`r coef_1` standard deviations. That estimate is about `r round(coef_1 / se_1, 1)` times its
standard error (`r se_1`) — much larger than chance alone would produce, so we take it as
evidence of a real (and because of randomization: causal) effect.*
`r unhide()`

```{r include = FALSE}
coef_2 <- coef(zols)["articleTreat_2"] %>% round(3) %>% unname()
se_2 <- vcov(zols) %>% diag()
se_2 <- se_2["articleTreat_2"] %>% sqrt() %>% round(3) %>% unname()
```

3. What is the causal effect of reading news about integration on supporting anti-discrimination policy? The causal effect is: `r num_fitb(coef_2)`. What is the estimated standard error of this $\hat{\beta}$? `r num_fitb(se_2)`. *What do these two statistics suggest about a causal effect?*

`r hide("Model answer")`
*The estimate (`r coef_2` standard deviations) is only about `r round(coef_2 / se_2, 1)`
times its standard error (`r se_2`) — a difference this small arises easily by chance, so
the experiment gives no convincing evidence that the integration article changed support.*
`r unhide()`

4. Make a coefficient plot to visualize these results. Use an AI like ChatGPT or Google Gemini to help you if you get stuck.
`r hide("R solution -> don't peek too early ;) !")`
```{r}
(plotdata <- zols %>% # Prepare data for plotting
  tidy() %>% # Convert regression results to a tidy data frame
  filter(term != "(Intercept)") %>% # Remove the intercept term
  mutate( # Rename treatment variables for clarity
    term = case_when(
      term == "articleTreat_1" ~ "Discrimination",
      term == "articleTreat_2" ~ "Acculturation")))

# Create the plot
ggplot(data = plotdata, aes(y = estimate, x = term)) +
  geom_hline(yintercept = 0, color = "orange",
             lty = "dashed") + # Add a horizontal line at y=0
  # Plot point estimates and confidence intervals
  geom_pointrange(aes(min = conf.low, max = conf.high)) +
  coord_flip() + # Flip coordinates for horizontal display
  labs(title = "Causal effect of news articles",
       x = "Article on",
       y = "Comparison to control group (article on Venus)
    in standard deviations",
    caption = "Note: Results are based on a weighted OLS with robust standard errors.
    N = 1085, and R2 = 0.012.") +
  theme_minimal() # Use a minimal theme for clean appearance
```
`r unhide()`
