---
title: "Exercise 2 (20 minutes)"
subtitle: "You practice: becoming the simulator — setting a confounder's strength and sign yourself, predicting the bias from the formula, and checking whether OLS agrees. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

num_fitb <- function(x, digits = 2, width = 6) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}
```

```{r hidden-setup, include = FALSE}
pacman::p_load(tidyverse, estimatr, modelsummary)
```

In exercise 1 you adjusted a real estimate — but you could never check whether the adjusted number
was *right*, because nobody knows the true effect of the Bali attack. Here you do know it,
**because you set it yourself.**

1. Copy in this simulator. It is the lecture's toy example, wrapped in a function so you can turn
   three knobs: the **true effect**, how strongly the confounder affects the **outcome**
   ($\beta_{C \rightarrow Y}$), and how strongly it drives **selection into treatment**
   ($\beta_{D \rightarrow C}$).

```{r sim-fn}
pacman::p_load(tidyverse, estimatr, modelsummary)

simulate_toy <- function(b_bali    = 1,     # THE TRUTH: causal effect of Bali
                         b_age     = 0.3,   # C -> Y : how much age drives xenophobia
                         selection = -0.1,  # C -> D : how age drives interview timing
                         seed      = 831983) {
  set.seed(seed)
  tibble(age = rnorm(n = 500, mean = 40, sd = 5)) %>%
    mutate(
      int_day = (selection * age) + rnorm(n = 500, mean = 0, sd = 1),
      bali    = if_else(int_day >= mean(int_day), 1, 0),
      xeno    = (b_bali * bali) + (b_age * age) + rnorm(n = 500, mean = 0, sd = 1)
    )
}

toydat <- simulate_toy() # The lecture's settings
```

`r hide("Stuck?")`
- Did `pacman::p_load(...)` run without red errors? Fix that first.
- Getting *different* numbers from the ones below? Check you kept `seed = 831983` and did not
  re-run `rnorm()` in between — the seed is what makes everyone's "random" data identical.
- Odd results? *Session → Restart R*, then rerun from the top.
- Still stuck? Ask in the [t-R-ouble forum](https://absalon.ku.dk/courses/70545/discussion_topics/563046?module_item_id=2037268) on Absalon.

`r unhide()`

```{r solutions, include = FALSE}
fit3 <- function(d) {
  bi <- coef(lm_robust(xeno ~ bali, data = d))["bali"]
  m  <- lm_robust(xeno ~ bali + age, data = d)
  list(bi = unname(bi),
       adj = unname(coef(m)["bali"]),
       cy = unname(coef(m)["age"]),
       dc = unname(coef(lm_robust(age ~ bali, data = d))["bali"]))
}
A <- fit3(simulate_toy())
B <- fit3(simulate_toy(b_age = 0.9))
C <- fit3(simulate_toy(b_age = 0))
D <- fit3(simulate_toy(selection = 0.1))
E <- fit3(simulate_toy(selection = 0))
```

2. **Verify the formula.** Fit the three regressions you need, and confirm that
   $\tilde\beta = \beta + (\beta_{C \rightarrow Y} \times \beta_{D \rightarrow C})$ actually holds.

`r hide("Hint")`
You need three models: `xeno ~ bali` (biased), `xeno ~ bali + age` (adjusted), and — the one
students forget — `age ~ bali`, which measures the *imbalance*.
`r unhide()`

`r hide("R solution — don't peek too early ;) !")`
```{r sol2}
ols_bi   <- lm_robust(xeno ~ bali, data = toydat)        # biased estimate
ols_mult <- lm_robust(xeno ~ bali + age, data = toydat)  # adjusted estimate
ols_dc   <- lm_robust(age ~ bali, data = toydat)         # the imbalance

b_bi <- coef(ols_bi)["bali"]
b_mu <- coef(ols_mult)["bali"]
b_cy <- coef(ols_mult)["age"]
b_dc <- coef(ols_dc)["bali"]

# Does the decomposition close?
c(bivariate = b_bi, adjusted_plus_OVB = b_mu + (b_cy * b_dc))
```
`r unhide()`

   The bivariate estimate is `r num_fitb(A$bi, digits = 2)`, the adjusted estimate is
   `r num_fitb(A$adj, digits = 2)`, and the omitted variable bias is
   `r num_fitb(A$cy * A$dc, digits = 2)`.
   Given that you *set* the true effect to **1**, which estimate got closer?
   `r mcq(c(answer = "The adjusted one — the bivariate estimate is badly biased downwards", "The bivariate one", "Both are equally close", "Neither is close"))`

3. **Predict before you compute.** Now make the confounder *three times stronger* on the outcome —
   `simulate_toy(b_age = 0.9)` — but change nothing else. What happens to the **bivariate**
   estimate?
   `r mcq(c("It gets closer to the truth of 1", "It stays about the same", answer = "It becomes even more biased — so much that it turns negative", "It doubles to about 2"))`

`r hide("R solution — check your prediction")`
```{r sol3, eval = FALSE}
toy_strong <- simulate_toy(b_age = 0.9)

lm_robust(xeno ~ bali, data = toy_strong)        # bivariate
lm_robust(xeno ~ bali + age, data = toy_strong)  # adjusted
```

The bivariate estimate is now about **`r round(B$bi, 2)`** — it has the **wrong sign**. A strong
enough confounder does not merely shrink an effect; it can make a genuinely positive effect look
negative. The adjusted estimate still lands near **`r round(B$adj, 2)`**.
`r unhide()`

4. **Now break one arrow.** Set `b_age = 0` — age no longer affects xenophobia at all — but leave
   `selection = -0.1`, so the two groups are *still* badly imbalanced on age.
   What is the omitted variable bias now?

   With `b_age = 0`, the bivariate estimate is `r num_fitb(C$bi, digits = 2)` and the
   imbalance $\beta_{D \rightarrow C}$ is still about `r num_fitb(C$dc, digits = 1, width = 5)` years.
   The bias is `r mcq(c(answer = "essentially zero", "still large and negative", "larger than before", "impossible to compute"))`.

`r hide("Why?")`
Because the bias is a **product**: $\beta_{C \rightarrow Y} \times \beta_{D \rightarrow C}$.
Multiply a big imbalance by zero and you get zero. **An imbalanced variable is only a confounder
if it also affects the outcome.** This is why "the groups differ on X" is not, by itself, an
argument that your estimate is biased — you have to make the case for *both* arrows.
`r unhide()`

5. **Flip the selection.** Go back to `b_age = 0.3`, but set `selection = 0.1` — now *older*
   people are more likely to end up in the treatment group. Predict the direction first, then run it.

   The bivariate estimate becomes `r num_fitb(D$bi, digits = 2)`, which is
   `r mcq(c(answer = "too large — the bias is now positive", "too small — the bias is still negative", "exactly right", "unchanged"))`

`r hide("R solution")`
```{r sol5, eval = FALSE}
toy_flip <- simulate_toy(selection = 0.1)

lm_robust(xeno ~ bali, data = toy_flip)
lm_robust(xeno ~ bali + age, data = toy_flip)
```

Both arrows are now positive ($+ \times + = +$), so the bias is positive and the raw estimate
**overstates** the true effect of 1. Same data-generating machinery, one sign flipped, and the
error changes direction — exactly as the sign table on the slides predicted.
`r unhide()`

6. **The last knob is the interesting one.** Set `selection = 0`: interview timing no longer
   depends on age at all. Fit *both* models and compare them.

`r hide("R solution — don't peek too early ;) !")`
```{r sol6}
toy_rand <- simulate_toy(selection = 0)

modelsummary(
  list("Bivariate" = lm_robust(xeno ~ bali, data = toy_rand),
       "Adjusted"  = lm_robust(xeno ~ bali + age, data = toy_rand)),
  coef_rename = c("bali" = "Exposed to Bali", "age" = "Age"),
  stars = TRUE, gof_map = c("nobs"),
  output = "kableExtra"
)
```
`r unhide()`

   **Write one sentence** as a `# comment` saying what is special about this case — then compare:

`r hide("Model answer")`
With `selection = 0` the two models give essentially the **same** treatment estimate
(≈ `r round(E$bi, 2)`), and both are close to the true effect of 1. Nothing is confounded, because
age no longer predicts who gets treated — the imbalance term $\beta_{D \rightarrow C}$ is
approximately zero, so the bias is zero no matter how strongly age affects xenophobia.

**This is what a randomized controlled trial looks like** (Lecture 6). Randomisation sets
$\beta_{D \rightarrow C}$ to zero *for every confounder at once* — measured or not — which is why
an RCT needs no controls, and why it outranks anything you can do with a regression.
`r unhide()`

7. **Bonus, for the fast.** Everything above kept `seed = 831983`. Re-run the default simulation
   with three or four different seeds and collect the adjusted estimate each time. How close to
   the true value of 1 does it get, and why is it never *exactly* 1?

`r hide("R solution")`
```{r sol7, eval = FALSE}
map_dbl(c(1, 42, 2024, 831983), function(s) {
  coef(lm_robust(xeno ~ bali + age, data = simulate_toy(seed = s)))["bali"]
})
```

The adjusted estimate scatters *around* 1 without ever landing exactly on it — that is
**sampling error** (Lecture 3), not bias. Bias means being systematically off in one direction;
sampling error means bouncing around the truth. Multiple OLS removed the bias here; only a larger
sample would shrink the bounce.
`r unhide()`

8. **Discuss with your neighbour.** In this exercise you always knew the truth. Work through:

   - In real research, which of the three quantities in the formula can you actually **estimate**
     from your data, and which would you have to argue for on theoretical grounds?
   - You adjust for a confounder and your estimate barely moves. Give **two** different
     explanations for that.
   - Someone claims: "I controlled for everything in the data set, so my estimate is causal."
     What is wrong with that sentence?

`r hide("Talking points")`
- For an **observed** confounder you can estimate both arrows — $\beta_{C \rightarrow Y}$ from the
  adjusted model and $\beta_{D \rightarrow C}$ from a balance table. For an **unobserved** one you
  can estimate neither; you can only reason about the plausible *sign* of each arrow and thus the
  likely direction of the bias. That reasoning is a normal, respectable part of a discussion section.
- Two explanations: (i) the variable genuinely was not much of a confounder — one of the two arrows
  is near zero, as in task 4; or (ii) it *is* a confounder, but its bias is being offset by another
  confounder pulling the opposite way. A stable coefficient is evidence, not proof.
- "Everything in the data set" is not "everything". The sentence quietly assumes the data set
  happens to contain every variable that affects both treatment and outcome — which is an
  assumption about the world, not a fact about the data, and it is untestable. The honest version
  is: "I adjusted for the confounders I could observe, and here is why I think the remaining ones
  are unlikely to overturn this."
`r unhide()`
