---
title: "Exercise 1 (20 minutes)"
subtitle: "You practice: adjusting a natural experiment for observed confounders — and reading what the adjustment did. When you're done, assist some of the others …"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)

num_fitb <- function(x, digits = 2, width = 6) {
  a <- format(round(x, digits), nsmall = digits)
  fitb(unique(c(a, sub(".", ",", a, fixed = TRUE))), width = width)
}
```

```{r hidden-read, include = FALSE}
pacman::p_load(tidyverse, haven, estimatr, modelsummary)
ESS_raw <- read_dta("../assets/Legewie_ESS_02.dta") # rendered from the repo
```

Last week you estimated the **raw** effect of the Bali attack in Portugal. Today you ask whether
that estimate survives once we hold the **observed** confounders constant.

1. Download `Legewie_ESS_02.dta` from Absalon, then copy in this starter code. It is last week's
   `prepare()` — extended to keep three background variables: **age**, **gender** and
   **employment status**.

```{r eval = FALSE}
pacman::p_load(tidyverse, haven, estimatr, modelsummary)
ESS_raw <- read_dta("Legewie_ESS_02.dta") # from your project folder
```

```{r prep}
event_date <- as.Date("2002-10-13")   # the Bali attack
win_begin  <- as.Date("2002-09-14")
win_end    <- as.Date("2002-10-20")

prepare <- function(data, countries) {
  data %>%
    mutate(
      int_date = as.Date(sprintf("%s-%s-%s", inwyr, inwmm, inwdd)),
      treat = case_when(
        int_date > event_date & int_date <= win_end   ~ "After the attack",
        int_date < event_date & int_date > win_begin  ~ "Before the attack",
        TRUE ~ NA_character_) %>% fct_relevel("Before the attack", "After the attack"),
      # Anti-immigrant index: average 7 items, z-standardise, flip so high = xenophobic
      anti_immi = rowMeans(across(c(imtcjob, imbleco, imbgeco, imueclt,
                                    imwbcnt, imwbcrm, imbghct)), na.rm = TRUE) %>%
        scale() %>% as.numeric(),
      anti_immi = max(anti_immi, na.rm = TRUE) - anti_immi,
      across(c(brncntr, mocntr, facntr), as_factor),
      # --- the three observed confounders --------------------------
      age  = inwyr - yrbrn,
      gndr = as_factor(gndr) %>% fct_drop(),
      empl_stat = case_when(
        pdwrk  == 1 ~ "Working",
        uempla == 1 ~ "Unemployed",
        rtrd   == 1 ~ "Retired",
        TRUE        ~ "Other"
      ) %>% factor() %>% fct_relevel("Working"),
      # -------------------------------------------------------------
      pspwght = pweight * dweight
    ) %>%
    filter(str_detect(cntry, countries) &
             brncntr == "yes" & mocntr == "yes" & facntr == "yes") %>%
    select(treat, anti_immi, cntry, pspwght, age, gndr, empl_stat) %>%
    drop_na()
}

ESS <- prepare(ESS_raw, "PT") # Portugal only
```

`r hide("Stuck?")`
- Did `pacman::p_load(...)` run without red errors? Fix that first.
- **`read_dta()` can't find the file?** `Legewie_ESS_02.dta` must sit **in your project folder** —
  the same folder as your `.Rproj`. Check with `list.files()`.
- Odd results? *Session → Restart R*, then rerun from the top.
- Still stuck? Ask in the [t-R-ouble forum](https://absalon.ku.dk/courses/70545/discussion_topics/563046?module_item_id=2037268) on Absalon.

`r unhide()`

```{r solutions, include = FALSE}
ols_bi  <- lm_robust(anti_immi ~ treat, data = ESS, weights = pspwght)
ols_adj <- lm_robust(anti_immi ~ treat + age + gndr + empl_stat,
                     data = ESS, weights = pspwght)

b_bi  <- coef(ols_bi)["treatAfter the attack"]
b_adj <- coef(ols_adj)["treatAfter the attack"]
se_adj <- sqrt(diag(vcov(ols_adj)))["treatAfter the attack"]
shift <- b_adj - b_bi
```

2. **Before you compute anything**, look at the balance table below. In Portugal, respondents
   interviewed *after* the attack are on average **older**. Xenophobia tends to *rise* with age.
   By the sign rule from the lecture, what should adjusting for age do to the estimate?
   `r mcq(c("Push it up — the bias is negative", answer = "Push it down — the bias is positive", "Leave it unchanged — age is balanced", "Impossible to say without running the model"))`

```{r balance-ex, echo = FALSE}
ESS %>%
  select(treat, age, anti_immi, pspwght) %>%
  rename(weights = pspwght) %>%
  datasummary_balance(~ treat, data = ., output = "kableExtra")
```

3. Now **write the line yourself.** Fit the **adjusted** model: regress `anti_immi` on `treat`
   *plus* `age`, `gndr` and `empl_stat`, weighted by `pspwght`. Put it side by side with the
   bivariate model.

`r hide("Hint")`
Add the controls with `+`, exactly as on the slides:
`lm_robust(anti_immi ~ treat + age + gndr + empl_stat, data = ESS, weights = pspwght)`
`r unhide()`

`r hide("R solution — don't peek too early ;) !")`
```{r sol3, eval = FALSE}
# The unadjusted comparison from last week
ols_bi <- lm_robust(anti_immi ~ treat, data = ESS, weights = pspwght)

# ... now holding the three observed confounders constant
ols_adj <- lm_robust(
  anti_immi ~ treat + age + gndr + empl_stat,
  data    = ESS,
  weights = pspwght
)

modelsummary(
  list("Bivariate" = ols_bi, "Adjusted" = ols_adj),
  coef_rename = c("treatAfter the attack" = "After the attack",
                  "age" = "Age", "gndrfemale" = "Female"),
  stars = TRUE, gof_map = c("nobs", "r.squared"),
  output = "kableExtra"
)
```

```{r sol3-show, ref.label = "sol3", echo = FALSE, eval = TRUE}
```
`r unhide()`

4. The **adjusted** effect of the Bali attack among the Portuguese is
   `r num_fitb(b_adj, digits = 2)` standard deviations, with a standard error of
   `r num_fitb(se_adj, digits = 2)`. Is it still statistically significant at the 5% level?
   `r mcq(c(answer = "Yes — the coefficient is more than twice its standard error", "No — the confidence interval now covers zero", "Cannot be determined from these numbers"))`

5. Compare the two columns. Adjusting moved the estimate by about
   `r num_fitb(shift, digits = 2)` standard deviations. **Write one sentence** as a `# comment`
   interpreting the *change* — then compare with the model answer.

`r hide("Model answer")`
Holding age, gender and employment status constant, respondents interviewed after the Bali attack
score about **0.28 standard deviations** higher on the anti-immigrant index than those interviewed
before — down slightly from the unadjusted **0.33**. The drop is what the omitted variable bias
formula predicted: the treated group was somewhat *older*, and age is (weakly) associated with more
xenophobia, so part of the raw gap was age, not the attack. The effect survives adjustment, which
makes it more credible — but it is **not** proof, because only *observed* confounders were closed.
`r unhide()`

6. Make a **coefficient plot** comparing the two estimates of the treatment effect, so a reader can
   see the adjustment at a glance. Drop the intercept and all the control coefficients.

`r hide("Hint")`
`bind_rows()` the two `tidy()` outputs with `.id = "model"`, then
`filter(term == "treatAfter the attack")`.
`r unhide()`

`r hide("R solution — don't peek too early ;) !")`
```{r sol6, out.width='90%', fig.height = 3, fig.width = 7, results = FALSE}
plotdata <- bind_rows(
  tidy(ols_bi),          # Model 1: unadjusted
  tidy(ols_adj),         # Model 2: adjusted
  .id = "model"
) %>%
  filter(term == "treatAfter the attack") %>%   # Keep only the treatment effect
  mutate(model = if_else(model == "1", "Bivariate", "Adjusted"))

ggplot(plotdata, aes(x = estimate, y = model)) +
  geom_vline(xintercept = 0, colour = "#901A1E", linetype = "dashed") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  labs(
    title = "The Bali effect in Portugal survives adjustment",
    x = "Difference in xenophobia vs. before the attack (SD)",
    y = ""
  ) +
  theme_minimal(base_size = 13)
```
`r unhide()`

7. **Bonus, for the fast.** Age was only one of three controls. Fit a model with **age alone**
   (`anti_immi ~ treat + age`) and compare it to the full adjusted model. Which of the three
   controls is doing most of the work?

`r hide("R solution")`
```{r sol7, eval = FALSE}
ols_age <- lm_robust(anti_immi ~ treat + age, data = ESS, weights = pspwght)

modelsummary(
  list("Bivariate" = ols_bi, "+ age" = ols_age, "+ age, gender, employment" = ols_adj),
  coef_rename = c("treatAfter the attack" = "After the attack"),
  coef_omit = "^(?!treat)", # Show only the treatment effect
  stars = TRUE, gof_map = c("nobs"),
  output = "kableExtra"
)
```

Age alone moves the estimate only from 0.33 to about **0.31** — the remaining movement comes from
gender and employment status. This is the product rule again: age is badly *imbalanced* here, but
it barely predicts xenophobia in Portugal, so its bias contribution is small.
`r unhide()`

8. **Discuss with your neighbour.** Our estimate barely moved when we adjusted for three observed
   confounders. Work through these three steps:

   - Does a *small* change when adding controls make you **more** or **less** confident that the
     remaining estimate is causal? Why?
   - Name one confounder that is **not** in this data set but might plausibly affect both when
     someone was interviewed *and* how xenophobic they are.
   - If such a variable exists, can any regression on this data set fix it?

`r hide("Talking points")`
- A small change is **mildly reassuring**, not proof. It shows the observed confounders were not
  driving the result. But it says nothing about unobserved ones — a stable coefficient is
  consistent both with "no confounding left" and with "the confounding that remains is invisible".
- Plausible unobserved candidates: whether someone lives in a rural area or a city (affects both
  how easily interviewers reach them *and* their exposure to immigration), household composition,
  media diet, or how politically engaged they are.
- **No.** Multiple OLS can only close backdoor paths through variables that are *in the data*.
  That is precisely why Legewie needed the as-if-random timing in the first place — the design,
  not the regression, is what carries most of the credibility here.
`r unhide()`
