---
title: "Exercise 1 (30) minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 3)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

```{tikz, DAG_3, echo = FALSE, out.width='50%'}
\usetikzlibrary{shapes,decorations,arrows,calc,arrows.meta,fit,positioning,quotes}
\tikzset{
  -Latex,auto,node distance =1 cm and 1 cm,semithick,
  state/.style ={ellipse, draw, minimum width = 0.7 cm},
  point/.style = {circle, draw, inner sep=0.04cm,fill,node contents={}},
  bidirected/.style={Latex-Latex,dashed},
  el/.style = {inner sep=2pt, align=left, sloped}
}

\begin{tikzpicture}
\sffamily
\node[state, align = center] (1) at (0,0) {Education};
\node[state, align = center] (2) [right = of 1] {\$};
\node[state, align = center] (3) [dashed, gray, above = of 1] {Parental \\ education};
\node[state, align = center] (4) [dashed, gray, below = of 1] {Cognitive \\ ability};
\node[state, align = center] (5) [dashed, gray, left = of 3] {$C?$};
\node[state, align = center] (6) [red, left = of 1] {Child labor \\ laws};

\path (1) edge (2);
\path (3) edge [dashed, gray] (2);
\path (4) edge [dashed, gray] (2);
\path (3) edge [dashed, gray] (1);
\path (4) edge [dashed, gray] (1);
\path (5) edge [dashed, gray] (1);
\path (5) edge [dashed, gray] (2);
\path (6) edge [red] (1);
\end{tikzpicture}
```

In 2000. Acemoglu and Angrist (2000) had an idea: Could child labor laws of the early 21st century be used as an instrument variable? They recognized that child labor laws in the different US federal states required that children had a minimum education before they were allowed to work. However, how many years of schooling that required differed across the various US federal states and also across time. They argued it was basically random in which year a certain federal state would reform its child labor laws. Thus by comparing persons who were subject to different child labor laws in different states and at different times, one could study the causal effect of compulsory education on wages.

```{r}
pacman::p_load(
  tidyverse, # Data manipulation,
  modelsummary, # for balance and regression tables.
  estimatr, # Regression for weighted data,
  modelr, # Turn results of lm() into a tibble,
  ivreg, # IV 2SLS,
  masteringmetrics) # Data and examples from Mastering Metrics

data("child_labor", package = "masteringmetrics")
```

1. The data set contains information from several US censuses (1950 to 1990). Importantly, it contains three categorical dummy variables for child labor law: `cl7`, `cl8`, and `cl9`. These three dummy variable identify people who lived in a US state where at the given time when they were children, the law required 7, 8, or 9 years of schooling before they were allowed to work. The data also contain information on respondents' US State of Birth `sob`, their year of birth `yob`, and the year in which the census data was collected `year`. Recode the three child labor law variables into on continuous predictor, and define the US state and the census year variables as categorical.

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
child_labor <- child_labor %>%
  mutate(
    # Generate continuous child labor law variable,
    chld_lb_lw = case_when(
      cl7 == 1 ~ 7, 
      cl8 == 1 ~ 8,
      cl9 == 1 ~ 9),
    # US State of birth as categorical,
    sob = factor(sob),
    # Census year as categorical.
    year = factor(year))
```
`r unhide()`

```{r q1, include = FALSE}
# Estimate the first stage
frst_stage <- lm_robust(indEduc ~ chld_lb_lw + yob + year + sob, data = child_labor)
# Regression table
modelsummary(list("1st Stage" = frst_stage), stars = TRUE,
             gof_map = c("nobs", "r.squared"))
```

```{r include = FALSE}
coef_frst <- coef(frst_stage)["chld_lb_lw"] %>% round(., 3) %>% as.scalar()
```

2. What is the first stage estimate of child labor laws, controlling for US state of birth, year of birth, and year of census? One more year of compulsory education translates into `r fitb(coef_frst, width = "5")` years of more education among US Americans.

`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label = "q1"}

```
`r unhide()`

```{r reduced, include = FALSE}
# Estimate the reduced stage
rdcd_stage <- lm_robust(lnwkwage ~ chld_lb_lw + yob + year + sob, data = child_labor)
# Regression table
modelsummary(list("Reduced Stage" = rdcd_stage), stars = TRUE,
             gof_map = c("nobs", "r.squared"))
```

```{r include = FALSE}
coef_rdcd <- coef(rdcd_stage)["chld_lb_lw"] %>% round(., 3) %>% as.scalar()
```
3. `lnwkwage` contains the logged wages. Estimate the reduced form using the same control variables. A year of compulsory education increases logged wages by: `r fitb(coef_rdcd, width = "5")`. According to Acemoglu and Angrist, this is `r mcq(c("Not a causal effect!", answer = "the causal effect of child labor law, if their assumption is true that differences between states and years are random", "The causal effect of education"))`.

`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label = "reduced"}

```
`r unhide()`

4. Use the Wald-IV estimator to get the LATE of education.
`r hide("R solution -> dont' peak to early ;) !")`
$\frac{`r coef_rdcd`}{`r coef_frst`}=`r coef_rdcd / coef_frst`$
`r unhide()`

5. Drop all observations from the US Federal States 4 and 8. Then get the predicted values after the first stage. That is, get that part of education that correlates with the child labor laws. Use the predicted values to explain logged wage using the same control variables as before. Compare the your Wald-IV estimate.
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
child_labor <- child_labor %>%
  filter(sob != 4 & sob != 8) %>%
  add_predictions(frst_stage) %>%
  rename(pred_indEduc = pred)

# Estimate the first stage
scnd_stage <- lm_robust(lnwkwage ~ pred_indEduc + yob + year + sob, data = child_labor)
# Regression table
modelsummary(list("2nd Stage" = scnd_stage), stars = TRUE,
             gof_map = c("nobs", "r.squared"))
```
`r unhide()`

6. Use `ivreg()` to perform all steps in one simple 2SLS estimation. Compare to your Wald-IV estimate and your second stage!
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
twosls <- ivreg(lnwkwage ~ indEduc + yob + year + sob | chld_lb_lw + yob + year + sob, data = child_labor)

# Regression table
modelsummary(list("2SLS" = twosls), stars = TRUE,
             gof_map = c("nobs", "r.squared"))
```
`r unhide()`