---
title: "Exercise 2 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 3)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

```{r include= FALSE}
pacman::p_load(
  tidyverse, # Data manipulation,
  haven, # Handle labelled data.
  ggplot2, # beautiful figures,
  estimatr, # Regression for weighted data,
  modelsummary) # regression tables with nice layout.

# Read the ESS round 9 data
ESS <- read_spss("../../../assets/ESS9e03_1.sav") %>%
  # Keep only Danes and Bulgarians
  filter(cntry == "DE" | cntry == "BG") %>% 
  mutate( # Define variables as categorical and continuous
    cntry = as_factor(cntry) %>% fct_relevel("Denmark"),
    gndr = as_factor(gndr),
    eduyrs = case_when( # Recode education to sensible levels
      eduyrs < 9 ~ 9,
      eduyrs > 21 ~ 21,
      TRUE ~zap_labels(eduyrs)),
    agea = zap_labels(agea),
    # Outcome: Subtract from maximum to turn the scale around
    imwbcnt = 10 - zap_labels(imwbcnt)) %>%
  # Keep only a minimum set of variables we need today,
  select(idno, cntry, pspwght, gndr, eduyrs, agea, imwbcnt) %>%
  drop_na() # Delete cases with missing values.
```

```{r ols1, include = FALSE}
ols_log <- lm_robust(imwbcnt ~ cntry + gndr + eduyrs + I(log2(agea)), data = ESS, weights = pspwght)

# Report results in a nicely-formatted table
modelsummary(list("Xeno" = ols_log),
             gof_map = c("nobs", "r.squared"), stars = TRUE)
```

```{r echo = FALSE}
coef <- coef(ols_log)["log2(agea)"] %>% round(., 3) %>% as.scalar()
se <- vcov(ols_log) %>% diag()
se <- se["log2(agea)"] %>% sqrt() %>% round(., 3) %>% as.scalar()
```
1. Let's continue with xenophobia and the ESS. Instead of a polynomial, use the $\log_2$ transformation. A doubling of age, is associated with an inclease in xenophobia of `r fitb(coef, width = "5")` scale points.

2. Which model fits better, the one with the polynomials or the logged one? According to the $R^2$ and $\text{Adj.}R^2$, `r mcq(c("The logged is better.", "The polynomial is better", answer = "They are equally good."))` 
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
ols_poly2 <- lm_robust(imwbcnt ~ cntry + gndr + eduyrs + poly(I(agea / 10), 2, raw = TRUE), 
                       data = ESS, weights = pspwght)

# Report results in a nicely-formatted table
modelsummary(list("Xeno" = ols_log, "Xeno" = ols_poly2),
             gof_map = c("nobs", "r.squared"), stars = TRUE)
```
`r unhide()`

3. Visualize the implied relationship between age and xenophobia.
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
(fict_dat <- tibble( # 1. Make fictional data
  # Realistic variation in x: Min to max
  agea = min(ESS$agea):max(ESS$agea), 
  # Hold other variables constant.
  gndr = "Female", # Women
  eduyrs = mean(ESS$eduyrs, na.rm = TRUE), # Max democracy
  cntry = "Germany")) # most recent date

(fict_dat <- predict( # 2. Get predictions
  ols_log, newdata = fict_dat,
  interval = "confidence", level = 0.95)$fit %>%
   as_tibble() %>% # Turn into a tibble, then
   bind_cols(fict_dat, .)) # Add to the synthetic data.

# 3. Plot the predictions
ggplot(data = fict_dat, aes(y = fit, x = agea)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.5) +
  geom_line() +
  labs(title = "Predictions based on multiple OLS",
       x = "By age in years
        (Gender, country, of residence, and education held constant)",
       y = "Predicted Xenophobia") +
  theme_minimal()
```
`r unhide()`



