---
title: "Exercise 2 (30 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(
  htmltools.dir.version = FALSE,
  servr.interval = 0.5,
  width = 115,
  digits = 3
  )
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

We're going to continue with our look at the "incumbent advantage" in US elections, again using the Lee2008 data that comes with RDDtools. But this time we are going to explore some of our methodological tools for getting a better estimate of the potential causal effect.

```{r results = FALSE}
pacman::p_load(
  tidyverse, # Data manipulation,
  ggplot2, # beautiful figures,
  texreg,# Regression tables with nice layout.
  devtools, # For installing RDDtools
  Rtools, # For installing RDDtools
  RDDtools # For running Regression Discontinuity analyses
)
```


```{r}
# Import the data from Lee, David. (2008). "Randomized experiments from non-random selection in U.S. House elections." Journal of Econometrics.
data('Lee2008')
```

```{r nonlinear_rdds, include=FALSE}
Lee2008_rdd <- RDDdata(
  y = Lee2008$y,
  x = Lee2008$x,
  cutpoint = 0
  )

reg_para1 <- RDDreg_lm(RDDobject = Lee2008_rdd, order = 1)
coef1 <- round(summary(reg_para1)$coefficients[2, 1], 3)
se1 <- round(summary(reg_para1)$coefficients[2, 2], 3)

reg_para2 <- RDDreg_lm(RDDobject = Lee2008_rdd, order = 2)
coef2 <- round(summary(reg_para2)$coefficients[2, 1], 3)
se2 <- round(summary(reg_para2)$coefficients[2, 2], 3)

reg_para3 <- RDDreg_lm(RDDobject = Lee2008_rdd, order = 3)
coef3 <- round(summary(reg_para3)$coefficients[2, 1], 3)
se3 <- round(summary(reg_para3)$coefficients[2, 2], 3)

reg_para4 <- RDDreg_lm(RDDobject = Lee2008_rdd, order = 4)
coef4 <- round(summary(reg_para4)$coefficients[2, 1], 3)
se4 <- round(summary(reg_para4)$coefficients[2, 2], 3)
```

1A. Fit separate models to the Lee2008 data where (a) you assume the relationship is linear, as in Exercise 1; (b) you assume the relationship is a second order polynomial, i.e. a power of two; and (c) you assume the relationship is a third order polynomial, i.e. a power of three, and (d) you assume the relationship is a fourth order polynomial, i.e. a power of four. Create a plot for each model, showing the line of best fit.

`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label="nonlinear_rdds"}

```
`r unhide()`

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
plot(
  reg_para1,
  xlab='Democratic Win Margin in time t\n',
  ylab='Democratic Vote Share in time t+1'
  )
```
`r unhide()`

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
plot(
  reg_para2,
  xlab='Democratic Win Margin in time t\n',
  ylab='Democratic Vote Share in time t+1'
  )
```
`r unhide()`

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
plot(
  reg_para3,
  xlab='Democratic Win Margin in time t\n',
  ylab='Democratic Vote Share in time t+1'
  )
```
`r unhide()`

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
plot(
  reg_para4,
  xlab='Democratic Win Margin in time t\n',
  ylab='Democratic Vote Share in time t+1'
  )
```
`r unhide()`

1B. Inspect each plot visually. Every time you increase the order of the polynomial the line appears to provide `r mcq(c("a worse", "the same", answer = "a better"))` fit for the data. Every model provides a different estimate for the effect of being the winner of the previous election. Among the models that assume a non-linear relationship, the largest estimate is (3 decimal places) `r fitb(coef3, width = "5")` and the smallest is (3 decimal places) `r fitb(coef2, width = "5")`.

`r hide("R solution Part 1 -> dont' peak to early ;) !")`
```{r}
summary(reg_para2)
```
`r unhide()`

`r hide("R solution Part 2")`
```{r}
summary(reg_para3)
```
`r unhide()`

`r hide("R solution Part 3")`
```{r}
summary(reg_para4)
```
`r unhide()`

```{r nonparam_rdd, include=FALSE}
bandwidth <- RDDbw_IK(Lee2008_rdd)
reg_nonparam <- RDDreg_np(RDDobject = Lee2008_rdd, bw = bandwidth)
coef_nonparam <- summary(reg_nonparam)$coefMat[1]
se_nonparam <- summary(reg_nonparam)$coefMat[2]

summary(reg_nonparam)
```

2. Now let's hone in on the comparison at the threshold. Fit a nonparametric model on the Lee2008 data. The estimated effect in the nonparametric model is (3 decimal places) `r fitb(coef_nonparam, width = "5")` and its standard error is (3 decimal places) `r fitb(se_nonparam, width = "5")`. This means that the effect is `r mcq(c("Not significant", answer = "Statistically significant", "worrysome"))`.

`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label="nonparam_rdd"}

```
`r unhide()`

3. Create a plot showing the results of your nonparametric RDD in Q2.

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
plot(
  reg_nonparam,
  xlab='Democratic Win Margin in time t\n',
  ylab='Democratic Vote Share in time t+1'
  )
```
`r unhide()`
