---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(
  htmltools.dir.version = FALSE,
  servr.interval = 0.5,
  width = 115,
  digits = 3
  )
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

We're going to look at the political advantage that comes with being an "incumbent" candidate in the United States. That is, does the winner of the previous election have higher chances of winning the following election as well. 

```{r results = FALSE}
pacman::p_load(
  tidyverse, # Data manipulation,
  ggplot2, # beautiful figures,
  texreg,# Regression tables with nice layout.
  devtools, # For installing RDDtools
  Rtools # For installing RDDtools on Windows, specifically
)
```

Getting the software running is a little more involved this time. We'll need to run this code to install the RDDtools package from GitHub. The good news is that RDDtools makes it *very* simple to produce RDD results.

```{r}
install_github("MatthieuStigler/RDDtools/RDDtools")
library(RDDtools)
```

The data come pre-packaged in RDDtools. Nice and easy.

```{r}
# Import the data from Lee, David. (2008). "Randomized experiments from non-random selection in U.S. House elections." Journal of Econometrics.
data('Lee2008')
```

Note that RDDtools will always convert your column names to "x" and "y", so there isn't much point in changing the column names directly.

In our case, x is the margin of victory for the Democratic candidate in the election that happened in time t (the previous election). When the number is positive, the Democratic candidate won. When the number is negative, the Democratic candidate lost. The number represents the proportion of the votes that was cast for the Democratic candidate, minus the proportion of the votes that was cast for the Republican candidate. If the election was 54% to 46% with the Democratic candidate winning, it would be entered as (0.54-0.46) 0.08 in the x column. Note: this means that the cutoff or threshold is 0 for this design.

Meanwhile, y is the proportion of the vote that the Democratic candidate wins in the election that happened in time t+1 (the following election). If the vote was again 54% to 46% for the Democratic candidate, the y column would hold 0.54.

1. Create a plot showing the discontinuity at x=0

`r hide("R solution -> don't peak too early ;) !")`
```{r}
Lee2008_rdd <- RDDdata(
  y = Lee2008$y,
  x = Lee2008$x,
  cutpoint = 0
  )

plot(
  Lee2008_rdd,
  xlab='Democratic Win Margin in time t\n',
  ylab='Democratic Vote Share in time t+1'
  )
```
`r unhide()`

```{r linear_rdd, include=FALSE}
reg_para <- RDDreg_lm(RDDobject = Lee2008_rdd, order = 1, slope='same')
coef <- round(summary(reg_para)$coefficients[2, 1], 3)
se <- round(summary(reg_para)$coefficients[2, 2], 3)

summary(reg_para)
```

2. Fit a parametric regression discontinuity model with the data, under the assumption that the relationship is linear and that the slopes are the same on both sides of the threshold. The estimated effect of winning the election in time t on vote shares in time t+1 is (3 decimal places) `r fitb(coef, width = "5")`. The standard error is `r fitb(se, width = "5")`, so the association is `r mcq(c("Not significant", answer = "Statistically significant", "worrysome"))`.

`r hide("R solution -> dont' peak to early ;) !")`
```{r ref.label='linear_rdd'}

```
`r unhide()`

3. Create a plot showing the results of your RDD in Q2.

`r hide("R solution -> don't peak too early ;) !")`
```{r}
plot(
  reg_para,
  xlab='Democratic Win Margin in time t\n',
  ylab='Democratic Vote Share in time t+1'
  )
```
`r unhide()`

