---
title: "Exercise 2 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

```{r}
pacman::p_load(
  tidyverse, # Data manipulation,
  ggplot2, # beautiful figures,
  estimatr, # Regression for weighted data,
  modelr, # Turn results of lm() into a tibble,
  modelsummary, # for balance and regression tables
  ivreg, # IV 2SLS,
  masteringmetrics) # Data and examples from Mastering Metrics

# Load the data from the Minneapolis Domestic Violence Experiment
data("mdve", package = "masteringmetrics")
```

1. The variable `T_RANDOM` contains the information about which strategy the police officers where supposed to use for the respective domestic violence emergency call: 1 indicates "Arrest", 2 "Advise to stop", 3 "Suspect Told to Leave", and 4 "Other". By contrast, the variable `T_FINAL` contains the information about what the police officers actually did when arriving at the scene. Recode the two variables into a respective factor variable, which combines 2 & 3 into a category "Coddle" versus 1 "Arrested", and 4 as `NA`. Make "Arrested" the reference. After recoding, create a cross-tabulation of the two variables.

How many police officers complied to the strategy that was *randomly* assigned to them? `r fitb(268, width = "3")`

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
mdve <- mdve %>%
 mutate( # Recode into factor variables with two categories
   T_RANDOM = case_when(
     T_RANDOM == 2 | T_RANDOM == 3 ~ "Coddle!", # Combine categories 2 and 3 into "Coddle!"
     T_RANDOM == 1 ~ "Arrest!", # Category 1 becomes "Arrest!"
     TRUE ~ as.character(NA)) %>% # All other values become NA
     fct_relevel("Arrest!"), # Make "Arrest!" the reference level
   T_FINAL = case_when(
     T_FINAL == 2 | T_FINAL == 3 ~ "Coddled", # Combine categories 2 and 3 into "Coddled"
     T_FINAL == 1 ~ "Arrested", # Category 1 becomes "Arrested"
     TRUE ~ as.character(NA)) %>% # All other values become NA
     fct_relevel("Arrested")) # Make "Arrested" the reference level

mdve %>%
 select(T_RANDOM, T_FINAL) %>% # Select only the treatment variables
 table() # Create a cross-tabulation of the two variables

# Compliers:
91 + 177
```
`r unhide()`

2. The variable `S_RACE` indicates the subject's (i.e. the person accused of domestic violence) race. The categories used by the police back in '81 where: 1 "White", 2 "Black", 3 "Indian", 4 "Asian", 5 "Hispanic", as well as sevel small other categories. Now make a balance table. Did the randomization work, or where certain groups more likely to face a police officers who got advised to arrest/coddle them? `r mcq(c("No, there is strong imbalance in the absolute frequencies", "Yes, the proprotions of the different groups are pretty similar across the two strategies", answer = "For 'Indians' the balance looks pretty off"))`.

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
mdve <- mdve %>%
 mutate( # Recode race variable
   S_RACE = case_when(
     S_RACE == 1 ~ "White", # Code 1 becomes "White"
     S_RACE == 2 ~ "Black", # Code 2 becomes "Black"
     S_RACE == 3 ~ "Indian", # Code 3 becomes "Indian"
     S_RACE == 4 ~ "Asian", # Code 4 becomes "Asian"
     S_RACE == 5 ~ "Hispanic", # Code 5 becomes "Hispanic"
     TRUE ~ "Other")) # All other values become "Other"

mdve %>%
 select(T_RANDOM, S_RACE) %>% # Select treatment and race variables for balance test
 datasummary_balance( # Create a balance table
   formula = ~ T_RANDOM, # Compare race distribution across treatment groups
   data = ., # Use the piped data
   title = "Subject's race by assigned policing strategy") # Add descriptive title
```
`r unhide()`

3. How about if we look at the strategy that the police officers actually used? `r mcq(c("There is still strong imbalance in the absolute frequencies", answer = "Interestingly, now the proprotions of the different groups are pretty similar across the two strategies", "Particularly white subjects got the more lenient treatment than randomly assigned."))`.

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
mdve %>%
 select(T_FINAL, S_RACE) %>% # Select treatment outcome and race variables for balance test
 datasummary_balance( # Create a balance table
   formula = ~ T_FINAL, # Compare race distribution across actual treatment groups
   data = ., # Use the piped data
   title = "Subject's race by assigned policing strategy") # Add descriptive title
```
`r unhide()`

```{r first_stage, include = FALSE}
mdve <- mdve %>%
 mutate(
   T_FINAL_01 = case_when(
     T_FINAL == "Arrested" ~ 0, # Code "Arrested" as 0
     T_FINAL == "Coddled" ~ 1)) # Code "Coddled" as 1

ols <- lm_robust(T_FINAL_01 ~ T_RANDOM, data = mdve) # Run robust linear regression of actual on assigned treatment

modelsummary(list("Repeat violence" = ols), # Create summary table of regression results
            stars = TRUE, # Add significance stars
            gof_map = c("nobs", "r.squared")) # Show only number of observations and R-squared as model fit statistics
```

```{r include = FALSE}
coef <- coef(ols, digits = 3)["T_RANDOMCoddle!"] %>% as.scalar()
```
4. How much higher was the probability that police officers would "coddle" a subject rather than arrest them, if they were randomly told to do so? `r fitb(coef, width = "3")` (three decimals). In IV language, this is called the `r mcq(c("Reduced form", answer = "First stage", "LATE estimate"))`.
`r hide("R solution -> dont' peak to early ;) !")` 
```{r ref.label = "first_stage"}
```
`r unhide()`

5. According to Sherman and Berk (1984, see also Angrist and Pischke's "Mastering Metrics"), the probability that a subject would repeat to assault their partner at home was **0.114** higher when the police officers were *randomly* told to coddle, as compared to when they were told to arrest (regardless of what they actually did at the scene). In IV language, this is called the `r mcq(c(answer = "Reduced form", "First stage", "LATE estimate"))`.

6. Based on the above information, what is the IV estimate of the local average causal effect of coddling subjects who are accused of domestic violence on the probability that they repeat to assault their partner? It is: `r fitb(0.14, width = "4")` (two decimals).
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
# The IV estimate is:
0.114  / 0.79
```
`r unhide()`