---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 3)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

Let's take another look at the Legewie, terrorist attack data. As before, run the code below to prepare the data for our analysis.

```{r results = FALSE}
pacman::p_load(
  tidyverse, # Data manipulation,
  haven, # Handle Stata data.
  ggplot2, # beautiful figures,
  estimatr, # Regression for weighted data,
  furniture, # For row-means,
  modelsummary) # regression tables with nice layout.

# Bali terror attack
event_date_begin <- as.Date("2002-09-14")
event_date       <- as.Date("2002-10-13")
event_date_end   <- as.Date("2002-10-20")

# Read the ESS 2002 data and prepare for analysis
ESS <- read_dta("../../../assets/Legewie_ESS_02.dta") %>%
  mutate(
    # Day and time of the survey interview,
    int_year = inwyr,
    int_mon  = inwmm,
    int_day  = inwdd,
    int_date = sprintf("%s-%s-%s", int_year, int_mon, int_day) %>% as.Date(),
    # Treatment indicator,
    treat = case_when(
      int_date > event_date & int_date <= event_date_end ~ "After the attack",
      int_date < event_date & int_date > event_date_begin ~ "Before the attack",
      TRUE ~ as.character(NA)) %>% fct_relevel("Before the attack", "After the attack"),
    # Attitudes towards immigrants
    anti_immi = rowmeans(
      imtcjob, imbleco, imbgeco, imueclt,
      imwbcnt, imwbcrm, imbghct, na.rm = TRUE) %>% scale() %>% as.numeric(),
    # Subtract from max to make larger values more xenophobic.
    anti_immi = max(anti_immi, na.rm = TRUE) - anti_immi,
    # Immigrant origin
    brncntr = as_factor(brncntr),
    mocntr = as_factor(mocntr),
    facntr = as_factor(facntr),
    #######################################
    # Age
    age = int_year - yrbrn, #<<
    # Gender
    gndr = as_factor(gndr) %>% fct_drop(),
    # Employment status
    empl_stat = case_when(
      pdwrk == 1 ~ "Working",
      uempla == 1 ~ "Unemployed",
      rtrd == 1 ~ "Retired",
      TRUE ~ "Other"),
    ########################################
    # Weights
    pspwght = pweight * dweight) %>%
  # Sample restriction        
  filter(cntry != "DK" & cntry != "IL" & cntry != "HU" & 
    # Mainstream members (i.e. immigrant origin)
    brncntr == "yes" & mocntr == "yes" &  facntr == "yes") %>%
  # Keep only a minimum set of variables we need today,
  select(treat, anti_immi, cntry, int_date, pspwght, age, empl_stat) %>%
  drop_na() # Delete cases with missing values.
```

1. This time, the data set contains interviews from people living in 10 different countries. Compared to Portugal as reference, are the terror effects significantly weaker or stronger in some countries? `r mcq(c("No they are not significantly different", answer = "Most are significantly smaller", "Most are significantly larger", "Only Slovenia differs"))`

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
ESS <- ESS %>%
  mutate( # Make Portugal the reference
    cntry = cntry %>% fct_relevel("PT"))

ols_inter <- lm_robust(anti_immi ~ treat*cntry, data = ESS, weights = pspwght)

# Report results in a nicely-formatted table
modelsummary(list("Xeno" = ols_inter),
             gof_map = c("nobs", "r.squared"), stars = TRUE)
```
`r unhide()`

```{r echo = FALSE}
coefPT <- coef(ols_inter)["treatAfter the attack"] %>% round(., 3) %>% as.scalar()
```
2. According to this regression model, what was the effect of the Bali terror attack in Portugal? `r fitb(coefPT, width = "5")`

```{r echo = FALSE}
coefSI <- coef(ols_inter)["treatAfter the attack"] + coef(ols_inter)["treatAfter the attack:cntrySI"] %>% round(., 3) %>% as.scalar()
```
3. According to this regression model, what was the effect of the Bali terror attack in Slovenia? `r fitb(coefSI, width = "5")`
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
coef(ols_inter)["treatAfter the attack"] + coef(ols_inter)["treatAfter the attack:cntrySI"]
```
`r unhide()`

4. Control for age and employment status. Are the results robust? `r mcq(c("No, there are several differences", answer = "Yes, and some countries are even not significantly different from Portugal anylonger"))`
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
ols_inter2 <- lm_robust(anti_immi ~ treat*cntry + age + empl_stat, data = ESS, weights = pspwght)

# Report results in a nicely-formatted table
modelsummary(list("Xeno" = ols_inter, "Xeno" = ols_inter2),
             gof_map = c("nobs", "r.squared"), stars = TRUE)
```
`r unhide()`
