---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

1. Let's take another look at our research question on the legacy of colonialism. Use the same setup as in lecture 4 and make Denmark the reference category. Then answer the following question: To which degree do differences in the number of years since independence from colonialism explain why we saw significant differences between countries colonized by different colonial powers? `r mcq(c("It does not explain much", "Many of the stars are gone.", answer = "The years since indepence explain a large share of the different poverty legacies of the various colonializers"))`. What does this say about a causal colonializer effect? `r mcq(c("It's much smaller than the bivariate result suggests.", "Correlation isn't causation", answer = "It is unclear whether the years since independence are not themselves an outcome of the colonializer and thus whether it is a good control if we are interested in a causal effect", "It's larger than the bivariate result suggests."))`
```{r}
pacman::p_load(
  tidyverse, # Data manipulation,
  ggplot2, # beautiful figures,
  wbstats, # download data from Worldbank. Tremendous source of global socio-economic data.
  estimatr, # OLS with robust SE,
  texreg) # regression tables with nice layout,

# 1. The worldbank data
Dat <- wb_data("SI.POV.DDAY", # Download poverty data: <1.90$ per day,
               start_date = 2002, end_date = 2022) %>%
  rename(poverty = SI.POV.DDAY) %>% # rename poverty variable,
  select(country, date, poverty) %>% # Keep only 3 variables
  drop_na(poverty) %>% # Drop cases with missing data,
  group_by(country) %>% # Group by country,
  filter(date == max(date)) %>% # Keep the most recent data per country.
  mutate(date = as.numeric(date)) %>% ungroup()

# 2. Colonizer data
Dat <- Dat %>% 
  mutate(
    colonizer = case_when(
      str_detect(country, "Algeria|Benin|Burkina Faso|Cambodia|Central African Republic|Chad") |
        str_detect(country, "Djibouti|Gabon|Guinea|Haiti|Laos|Lebanon|Madagascar|Mali|Congo, Rep.|Cote D'Ivoire") | 
        str_detect(country, "Mauritania|Niger|Senegal|Syria|Togo|Tunisia|Vietnam") ~ "France", # France
      str_detect(country, "Angola|Brazil|Mozambique|Guinea-Bissau") ~ "Portugal", # Portugal
      str_detect(country, "Argentina|Bolivia|Chile|Colombia|Costa Rica|Cuba|Dominican Republic|Ecuador") |
        str_detect(country, "El Salvador|Guatemala|Honduras|Mexico|Nicaragua|Panama|Paraguay|Peru") |
        str_detect(country, "Uruguay|Venezuela") ~ "Spain", # Spain
      str_detect(country, "Australia|Bahrain|Bangladesh|Botswana|Canada|Cyprus|Egypt|Fiji") |
        str_detect(country, "Gambia|Ghana|Guyana|India|Iraq|Israel|Jamaica|Jordan|Kenya|Kuwait") |
        str_detect(country, "Lesotho|Malawi|Malaysia|Mauritius|Myanmar|New Zealand|Nigeria|Oman") | 
        str_detect(country, "Pakistan|Qatar|Sierra Leone|Singapore|South Africa") | 
        str_detect(country, "Sri Lanka|Sudan|Swaziland|Tanzania|Trinidad/Tobago|Uganda") | 
        str_detect(country, "United Arab Emirates|United States|Yemen|Zambia|Zimbabwe") ~ "Britain", # Britain
      str_detect(country, "Burundi|Congo, Dem. Rep.|Rwanda") ~ "Belgium",
    str_detect(country, "Iceland|Estonia") ~ "Denmark")) # Denmark

# 3. Independence from colonialism data
Dat <- Dat %>% 
  mutate(
    years_indep = case_when(
      is.na(colonizer) ~ as.numeric(NA),
      country == "Estonia" ~ date - 1645,
      country == "United States" ~ date - 1783, country == "Haiti" ~ date - 1804, 
      country == "Paraguay" ~ date - 1811, country == "Chile" ~ date - 1818,
      str_detect(country, "Argentina|Bolivia|Colombia") ~ date - 1819,
      str_detect(country, "Costa Rica|Dominican Republic|Mexico|Nicaragua|Panama|El Salvador|Guatemala|Honduras|Venezuela") ~ date - 1821,
      str_detect(country, "Brazil|Ecuador") ~ date - 1822,
      country == "Peru" ~ date - 1824, country == "Uruguay" ~ date - 1828,
      country == "Cuba" ~ date - 1899, country == "Australia" ~ date - 1901,
      country == "New Zealand" ~ date - 1907, country == "South Africa" ~ date - 1910,
      country == "Egypt" ~ date - 1922, country == "Iraq" ~ date - 1932,
      str_detect(country, "Iceland") ~ date - 1944,
      str_detect(country, "Korea|Taiwan|Vietnam") ~ date - 1945,
      str_detect(country, "Lebanon|Philippines|Syria") ~ date - 1946,
      str_detect(country, "Bangladesh|Pakistan|India|Liberia") ~ date - 1947,
      str_detect(country, "Myanmar|Israel|Jordan|Sri Lanka") ~ date - 1948,
      country == "Indonesia" ~ date - 1949, country == "Libya" ~ date - 1951,
      str_detect(country, "Cambodia|Loas") ~ date - 1954,
      str_detect(country, "Morocco|Sudan|Tunisia") ~ date - 1956,
      str_detect(country, "Malaysia|Ghana") ~ date - 1957,
      country == "Guinea" ~ date - 1958, country == "Singapore" ~ date - 1959,
      str_detect(country, "Benin|Burkina Faso|Central African Republic|Chad|Congo, Dem. Rep.|Congo, Rep.|Code D'Ivoire|Mali|Mauritania|Niger|Nigeria|Senegal|Gabon|Somalia|Togo") ~ date - 1960,
      str_detect(country, "Kuwait|Sierra Leone|Tanzania") ~ date - 1961,
      str_detect(country, "Algeria|Burundi|Rwanda|Jamaica|Trinidad/ Tobago|Uganda") ~ date - 1962,
      country == "Kenya" ~ date - 1963,
      str_detect(country, "Malawi|Zambia") ~ date - 1964,
      str_detect(country, "Gambia|Zimbabwe") ~ date - 1965,
      str_detect(country, "Botswana|Lesotho|Guyana") ~ date - 1966,
      str_detect(country, "Canada|Yemen") ~ date - 1967,
      str_detect(country, "Mauritius|Swaziland") ~ date - 1968,
      country == "Fiji" ~ date - 1970,
      str_detect(country, "Bahrain|Oman|Qatar|United Arab Emirates") ~ date - 1971,
      country == "Guinea-Bissau" ~ date - 1974,
      str_detect(country, "Angola|Mozambigue|Papua New Guinea") ~ date - 1975,
      country == "Djibouti" ~ date - 1977, 
      country == "Namibia" ~ date - 1990))
```

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
Dat <- Dat %>%
  mutate( # Make Denmark the reference
    colonizer = colonizer %>% fct_relevel("Denmark"))

# Bivariate model
ols_bi <- lm_robust(poverty ~  colonizer, data = Dat)
# Controlling for years since independence
ols_mult <- lm_robust(poverty ~  colonizer + years_indep, data = Dat)

screenreg(list(ols_bi, ols_mult), # regression table.
          include.ci = FALSE, digits = 3)
```
`r unhide()`

2. Make a coefficient plot that visualizes the country differences (compared to Denmark as reference) before and after adjusting for the years since independence.
`r hide("R solution -> dont' peak to early ;) !")`
```{r Coefplot2, out.width='100%', fig.height = 4, fig.width = 6, eval = FALSE}
plotdata_1 <- ols_bi %>%
  tidy() %>% # Turn results into a tibble,
  mutate( # Rename variables for the plot.
    term = case_when(
      term == "colonizerBelgium" ~ "Belgium",
      term == "colonizerBritain" ~ "Britain",
      term == "colonizerFrance" ~ "France",
      term == "colonizerPortugal" ~ "Portugal",
      term == "colonizerSpain" ~ "Spain")) %>%
      filter(term != "(Intercept)")

plotdata_2 <- ols_mult %>%
  tidy() %>% # Turn results into a tibble,
  mutate( # Rename variables for the plot.
    term = case_when(
      term == "colonizerBelgium" ~ "Belgium",
      term == "colonizerBritain" ~ "Britain",
      term == "colonizerFrance" ~ "France",
      term == "colonizerPortugal" ~ "Portugal",
      term == "colonizerSpain" ~ "Spain")) %>%
      filter(term != "(Intercept)" & term != "years_indep")

plotdata <- bind_rows(
  plotdata_1, plotdata_2, .id = "model") %>%
  mutate(
    model = case_when(
      model == 1 ~ "Bivariate",
      model == 2 ~ "Adjusted for years since independence"))

ggplot(data = plotdata, 
       aes(y = estimate, 
           # Order by effect size
           x = reorder(term, estimate),
           color = model)) +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +
  # Point with error-bars,
  geom_pointrange(aes(min = conf.low, max = conf.high),
                  position = position_dodge(width=0.5)) +
  coord_flip() + # Flip Y- & X-Axis,
  labs(
    title = "Multiple OLS of extreme poverty",
    x = "",
    color = "",
    # Write Greek beta into axis title.
    y = "Difference in extreme poverty 
    compared to countries colonized by Denmark") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
`r unhide()`
```{r ref.label = "Coefplot2", out.width='100%', fig.height = 4, fig.width = 6, echo = FALSE}
```

3. Visualize model predictions that show the association of years since independence (from 0 to 300 years) and extreme poverty controlled for colonializer.
`r hide("R solution -> dont' peak to early ;) !")`
```{r predplot2, out.width='100%', fig.height = 4, fig.width = 6, eval = FALSE}
# Make fictional data
fict_dat <- tibble(
  years_indep = 0:300,
  colonizer = "Denmark")

# Get predictions
fict_dat <- predict(
  ols_mult, newdata = fict_dat,
  interval = "confidence", level = 0.95)$fit %>%
  as_tibble() %>% # Turn into a tibble, then
  bind_cols(fict_dat, .) # Add to the synthetic data.

ggplot(data = fict_dat, aes(y = fit, x = years_indep)) +
geom_vline(xintercept = c(34, 236), color = "red", lty = "dashed") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.5) +
  geom_line() +
  labs(
    title = "Predictions based on multiple OLS of extreme poverty",
    x = "By years since independence from colonialism
    (Colonializer set to: Denmark)",
    # Write Greek beta into axis title.
    y = "Predicted average of extreme poverty",
    caption = "") +
  theme_minimal()
```
`r unhide()`
```{r ref.label = "predplot2", out.width='100%', fig.height = 4, fig.width = 6, echo = FALSE}
```