---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

1. Let's take another look at our research question on the legacy of colonialism. Use the setup below.

```{r}
pacman::p_load(
  tidyverse, # Data manipulation,
  ggplot2, # beautiful figures,
  wbstats, # download data from Worldbank. Tremendous source of global socio-economic data.
  estimatr, # OLS with robust SE,
  modelsummary) # for balance and regression tables.

# 1. The worldbank data
Dat <- wb_data("SI.POV.DDAY", # Download poverty data: <1.90$ per day,
               start_date = 2002, end_date = 2022) %>%
  rename(poverty = SI.POV.DDAY) %>% # rename poverty variable,
  select(country, date, poverty) %>% # Keep only 3 variables
  drop_na(poverty) %>% # Drop cases with missing data,
  group_by(country) %>% # Group by country,
  filter(date == max(date)) %>% # Keep the most recent data per country.
  mutate(date = as.numeric(date)) %>% ungroup()

# 2. Colonizer data
Dat <- Dat %>% 
  mutate(
    colonizer = case_when(
      str_detect(country, "Algeria|Benin|Burkina Faso|Cambodia|Central African Republic|Chad") |
        str_detect(country, "Djibouti|Gabon|Guinea|Haiti|Laos|Lebanon|Madagascar|Mali|Congo, Rep.|Cote D'Ivoire") | 
        str_detect(country, "Mauritania|Niger|Senegal|Syria|Togo|Tunisia|Vietnam") ~ "France", # France
      str_detect(country, "Angola|Brazil|Mozambique|Guinea-Bissau") ~ "Portugal", # Portugal
      str_detect(country, "Argentina|Bolivia|Chile|Colombia|Costa Rica|Cuba|Dominican Republic|Ecuador") |
        str_detect(country, "El Salvador|Guatemala|Honduras|Mexico|Nicaragua|Panama|Paraguay|Peru") |
        str_detect(country, "Uruguay|Venezuela") ~ "Spain", # Spain
      str_detect(country, "Australia|Bahrain|Bangladesh|Botswana|Canada|Cyprus|Egypt|Fiji") |
        str_detect(country, "Gambia|Ghana|Guyana|India|Iraq|Israel|Jamaica|Jordan|Kenya|Kuwait") |
        str_detect(country, "Lesotho|Malawi|Malaysia|Mauritius|Myanmar|New Zealand|Nigeria|Oman") | 
        str_detect(country, "Pakistan|Qatar|Sierra Leone|Singapore|South Africa") | 
        str_detect(country, "Sri Lanka|Sudan|Swaziland|Tanzania|Trinidad/Tobago|Uganda") | 
        str_detect(country, "United Arab Emirates|United States|Yemen|Zambia|Zimbabwe") ~ "Britain", # Britain
      str_detect(country, "Burundi|Congo, Dem. Rep.|Rwanda") ~ "Belgium",
    str_detect(country, "Iceland|Estonia") ~ "Denmark")) # Denmark

# 3. Independence from colonialism data
Dat <- Dat %>% 
  mutate(
    years_indep = case_when(
      is.na(colonizer) ~ as.numeric(NA),
      country == "Estonia" ~ date - 1645,
      country == "United States" ~ date - 1783, country == "Haiti" ~ date - 1804, 
      country == "Paraguay" ~ date - 1811, country == "Chile" ~ date - 1818,
      str_detect(country, "Argentina|Bolivia|Colombia") ~ date - 1819,
      str_detect(country, "Costa Rica|Dominican Republic|Mexico|Nicaragua|Panama|El Salvador|Guatemala|Honduras|Venezuela") ~ date - 1821,
      str_detect(country, "Brazil|Ecuador") ~ date - 1822,
      country == "Peru" ~ date - 1824, country == "Uruguay" ~ date - 1828,
      country == "Cuba" ~ date - 1899, country == "Australia" ~ date - 1901,
      country == "New Zealand" ~ date - 1907, country == "South Africa" ~ date - 1910,
      country == "Egypt" ~ date - 1922, country == "Iraq" ~ date - 1932,
      str_detect(country, "Iceland") ~ date - 1944,
      str_detect(country, "Korea|Taiwan|Vietnam") ~ date - 1945,
      str_detect(country, "Lebanon|Philippines|Syria") ~ date - 1946,
      str_detect(country, "Bangladesh|Pakistan|India|Liberia") ~ date - 1947,
      str_detect(country, "Myanmar|Israel|Jordan|Sri Lanka") ~ date - 1948,
      country == "Indonesia" ~ date - 1949, country == "Libya" ~ date - 1951,
      str_detect(country, "Cambodia|Loas") ~ date - 1954,
      str_detect(country, "Morocco|Sudan|Tunisia") ~ date - 1956,
      str_detect(country, "Malaysia|Ghana") ~ date - 1957,
      country == "Guinea" ~ date - 1958, country == "Singapore" ~ date - 1959,
      str_detect(country, "Benin|Burkina Faso|Central African Republic|Chad|Congo, Dem. Rep.|Congo, Rep.|Code D'Ivoire|Mali|Mauritania|Niger|Nigeria|Senegal|Gabon|Somalia|Togo") ~ date - 1960,
      str_detect(country, "Kuwait|Sierra Leone|Tanzania") ~ date - 1961,
      str_detect(country, "Algeria|Burundi|Rwanda|Jamaica|Trinidad/ Tobago|Uganda") ~ date - 1962,
      country == "Kenya" ~ date - 1963,
      str_detect(country, "Malawi|Zambia") ~ date - 1964,
      str_detect(country, "Gambia|Zimbabwe") ~ date - 1965,
      str_detect(country, "Botswana|Lesotho|Guyana") ~ date - 1966,
      str_detect(country, "Canada|Yemen") ~ date - 1967,
      str_detect(country, "Mauritius|Swaziland") ~ date - 1968,
      country == "Fiji" ~ date - 1970,
      str_detect(country, "Bahrain|Oman|Qatar|United Arab Emirates") ~ date - 1971,
      country == "Guinea-Bissau" ~ date - 1974,
      str_detect(country, "Angola|Mozambigue|Papua New Guinea") ~ date - 1975,
      country == "Djibouti" ~ date - 1977, 
      country == "Namibia" ~ date - 1990))
```

2. Make Denmark the reference category. 
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
Dat <- Dat %>%
  mutate( # Make Denmark the reference
    colonizer = colonizer %>% fct_relevel("Denmark"))
```
`r unhide()`

3. Answer the following question: To which degree do differences in the number of years since independence from colonialism explain why we saw significant differences between countries colonized by different colonial powers? `r mcq(c("It does not explain much", "Many of the stars are gone.", answer = "The years since indepence explain a large share of the different poverty legacies of the various colonializers"))`. 

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
# Bivariate model
ols_bi <- lm_robust(poverty ~  colonizer, data = Dat)
# Controlling for years since independence
ols_mult <- lm_robust(poverty ~  colonizer + years_indep, data = Dat)

# Regression table
modelsummary(list("Biv. OLS" = ols_bi, "Mult. OLS" = ols_mult), 
             gof_map = c("nobs", "r.squared"), stars = TRUE)
```
`r unhide()`

4. What does this say about a causal colonializer effect? `r mcq(c("It's much smaller than the bivariate result suggests.", "Correlation isn't causation", answer = "It's unclear whether the years since independence are not themselves an outcome of the colonializer (i.e. a mediator). Thus, it might be bad control if we are interested in a causal effect of the colonializer", "It's larger than the bivariate result suggests."))`