---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

1. Continue with our research question on the legacy of colonialism. Try to replicate the coefficient plot shown below, which visualizes the country differences (compared to Denmark as reference) before and after adjusting for the years since independence. Use AI to understand how to use `position_dodge()` to produce this figure.

```{r echo = FALSE, results='hide'}
pacman::p_load(
  tidyverse, # Data manipulation,
  ggplot2, # beautiful figures,
  wbstats, # download data from Worldbank. Tremendous source of global socio-economic data.
  estimatr, # OLS with robust SE,
  modelsummary) # for balance and regression tables.

# 1. The worldbank data
Dat <- wb_data("SI.POV.DDAY", # Download poverty data: <1.90$ per day,
               start_date = 2002, end_date = 2022) %>%
  rename(poverty = SI.POV.DDAY) %>% # rename poverty variable,
  select(country, date, poverty) %>% # Keep only 3 variables
  drop_na(poverty) %>% # Drop cases with missing data,
  group_by(country) %>% # Group by country,
  filter(date == max(date)) %>% # Keep the most recent data per country.
  mutate(date = as.numeric(date)) %>% ungroup()

# 2. Colonizer data
Dat <- Dat %>% 
  mutate(
    colonizer = case_when(
      str_detect(country, "Algeria|Benin|Burkina Faso|Cambodia|Central African Republic|Chad") |
        str_detect(country, "Djibouti|Gabon|Guinea|Haiti|Laos|Lebanon|Madagascar|Mali|Congo, Rep.|Cote D'Ivoire") | 
        str_detect(country, "Mauritania|Niger|Senegal|Syria|Togo|Tunisia|Vietnam") ~ "France", # France
      str_detect(country, "Angola|Brazil|Mozambique|Guinea-Bissau") ~ "Portugal", # Portugal
      str_detect(country, "Argentina|Bolivia|Chile|Colombia|Costa Rica|Cuba|Dominican Republic|Ecuador") |
        str_detect(country, "El Salvador|Guatemala|Honduras|Mexico|Nicaragua|Panama|Paraguay|Peru") |
        str_detect(country, "Uruguay|Venezuela") ~ "Spain", # Spain
      str_detect(country, "Australia|Bahrain|Bangladesh|Botswana|Canada|Cyprus|Egypt|Fiji") |
        str_detect(country, "Gambia|Ghana|Guyana|India|Iraq|Israel|Jamaica|Jordan|Kenya|Kuwait") |
        str_detect(country, "Lesotho|Malawi|Malaysia|Mauritius|Myanmar|New Zealand|Nigeria|Oman") | 
        str_detect(country, "Pakistan|Qatar|Sierra Leone|Singapore|South Africa") | 
        str_detect(country, "Sri Lanka|Sudan|Swaziland|Tanzania|Trinidad/Tobago|Uganda") | 
        str_detect(country, "United Arab Emirates|United States|Yemen|Zambia|Zimbabwe") ~ "Britain", # Britain
      str_detect(country, "Burundi|Congo, Dem. Rep.|Rwanda") ~ "Belgium",
    str_detect(country, "Iceland|Estonia") ~ "Denmark")) # Denmark

# 3. Independence from colonialism data
Dat <- Dat %>% 
  mutate(
    years_indep = case_when(
      is.na(colonizer) ~ as.numeric(NA),
      country == "Estonia" ~ date - 1645,
      country == "United States" ~ date - 1783, country == "Haiti" ~ date - 1804, 
      country == "Paraguay" ~ date - 1811, country == "Chile" ~ date - 1818,
      str_detect(country, "Argentina|Bolivia|Colombia") ~ date - 1819,
      str_detect(country, "Costa Rica|Dominican Republic|Mexico|Nicaragua|Panama|El Salvador|Guatemala|Honduras|Venezuela") ~ date - 1821,
      str_detect(country, "Brazil|Ecuador") ~ date - 1822,
      country == "Peru" ~ date - 1824, country == "Uruguay" ~ date - 1828,
      country == "Cuba" ~ date - 1899, country == "Australia" ~ date - 1901,
      country == "New Zealand" ~ date - 1907, country == "South Africa" ~ date - 1910,
      country == "Egypt" ~ date - 1922, country == "Iraq" ~ date - 1932,
      str_detect(country, "Iceland") ~ date - 1944,
      str_detect(country, "Korea|Taiwan|Vietnam") ~ date - 1945,
      str_detect(country, "Lebanon|Philippines|Syria") ~ date - 1946,
      str_detect(country, "Bangladesh|Pakistan|India|Liberia") ~ date - 1947,
      str_detect(country, "Myanmar|Israel|Jordan|Sri Lanka") ~ date - 1948,
      country == "Indonesia" ~ date - 1949, country == "Libya" ~ date - 1951,
      str_detect(country, "Cambodia|Loas") ~ date - 1954,
      str_detect(country, "Morocco|Sudan|Tunisia") ~ date - 1956,
      str_detect(country, "Malaysia|Ghana") ~ date - 1957,
      country == "Guinea" ~ date - 1958, country == "Singapore" ~ date - 1959,
      str_detect(country, "Benin|Burkina Faso|Central African Republic|Chad|Congo, Dem. Rep.|Congo, Rep.|Code D'Ivoire|Mali|Mauritania|Niger|Nigeria|Senegal|Gabon|Somalia|Togo") ~ date - 1960,
      str_detect(country, "Kuwait|Sierra Leone|Tanzania") ~ date - 1961,
      str_detect(country, "Algeria|Burundi|Rwanda|Jamaica|Trinidad/ Tobago|Uganda") ~ date - 1962,
      country == "Kenya" ~ date - 1963,
      str_detect(country, "Malawi|Zambia") ~ date - 1964,
      str_detect(country, "Gambia|Zimbabwe") ~ date - 1965,
      str_detect(country, "Botswana|Lesotho|Guyana") ~ date - 1966,
      str_detect(country, "Canada|Yemen") ~ date - 1967,
      str_detect(country, "Mauritius|Swaziland") ~ date - 1968,
      country == "Fiji" ~ date - 1970,
      str_detect(country, "Bahrain|Oman|Qatar|United Arab Emirates") ~ date - 1971,
      country == "Guinea-Bissau" ~ date - 1974,
      str_detect(country, "Angola|Mozambigue|Papua New Guinea") ~ date - 1975,
      country == "Djibouti" ~ date - 1977, 
      country == "Namibia" ~ date - 1990))

Dat <- Dat %>%
  mutate( # Make Denmark the reference
    colonizer = colonizer %>% fct_relevel("Denmark"))

# Bivariate model
ols_bi <- lm_robust(poverty ~  colonizer, data = Dat)
# Controlling for years since independence
ols_mult <- lm_robust(poverty ~  colonizer + years_indep, data = Dat)

# Regression table
modelsummary(list("Biv. OLS" = ols_bi, "Mult. OLS" = ols_mult), 
             gof_map = c("nobs", "r.squared"), stars = TRUE)
```

`r hide("R solution -> dont' peak to early ;) !")`
```{r Coefplot2, out.width='100%', fig.height = 4, fig.width = 6, eval = FALSE}
plotdata_1 <- ols_bi %>%
  tidy() %>%  # Turn the regression results into a tidy tibble
  mutate(  # Rename the 'term' variable for better plot labels
    term = case_when(
      term == "colonizerBelgium" ~ "Belgium",  # If the term is "colonizerBelgium", change it to "Belgium"
      term == "colonizerBritain" ~ "Britain",  # And so on for other colonizers
      term == "colonizerFrance" ~ "France",
      term == "colonizerPortugal" ~ "Portugal",
      term == "colonizerSpain" ~ "Spain")) %>%
  filter(term != "(Intercept)")  # Remove the intercept term from the plot data

plotdata_2 <- ols_mult %>%
  tidy() %>%  # Turn the regression results into a tidy tibble
  mutate(  # Rename the 'term' variable for better plot labels (same as above)
    term = case_when(
      term == "colonizerBelgium" ~ "Belgium",
      term == "colonizerBritain" ~ "Britain",
      term == "colonizerFrance" ~ "France",
      term == "colonizerPortugal" ~ "Portugal",
      term == "colonizerSpain" ~ "Spain")) %>%
  filter(term != "(Intercept)" & term != "years_indep")  # Remove the intercept and 'years_indep' terms

plotdata <- bind_rows(  # Combine the two plot datasets
  plotdata_1, plotdata_2, .id = "model") %>%  # Add an identifier column called 'model'
  mutate(
    model = case_when(  # Change the 'model' identifier to more descriptive labels
      model == 1 ~ "Bivariate",  # If model == 1, label it "Bivariate"
      model == 2 ~ "Adjusted for years since independence"))  # If model == 2, label it "Adjusted for years since independence"

ggplot(data = plotdata, 
       aes(y = estimate,  # Set the y-axis to the estimate
           x = reorder(term, estimate),  # Set the x-axis to the term, reordered by the estimate value
           color = model)) +  # Color the points by the model type
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +  # Add a horizontal dashed line at y = 0
  geom_pointrange(aes(min = conf.low, max = conf.high),  # Add points with error bars representing confidence intervals
                  position = position_dodge(width=0.5)) +  # Dodge the points slightly to avoid overlap
  coord_flip() +  # Flip the x and y axes
  labs(
    title = "Multiple OLS of extreme poverty",  # Set the plot title
    x = "",  # Remove the x-axis label
    color = "",  # Remove the legend title
    y = "Difference in extreme poverty compared to countries colonized by Denmark") +  # Set the y-axis label
  theme_minimal() +  # Use a minimal theme
  theme(legend.position = "bottom")  # Place the legend at the bottom of the plot
```
`r unhide()`
```{r ref.label = "Coefplot2", out.width='100%', fig.height = 4, fig.width = 6, echo = FALSE}
```

2. This one is demanding! Try to replicate the figure shown below, which visualizes model predictions that show the association of years since independence (from 0 to 300 years) and extreme poverty controlled for colonializer. Also try to mark the minimum (34) and maximum (236) years of independence.
`r hide("R solution -> dont' peak to early ;) !")`
```{r predplot2, out.width='100%', fig.height = 4, fig.width = 6, eval = FALSE}
# Make fictional data
fict_dat <- tibble(  # Create a tibble for fictional data
  years_indep = 0:300,  # Generate a sequence of numbers from 0 to 300 representing years since independence
  colonizer = "Denmark")  # Set the colonizer to "Denmark" for all observations

# Get predictions
fict_dat <- predict(  # Predict values using the 'ols_mult' model (multiple linear regression)
  ols_mult, newdata = fict_dat,  # Use the 'fict_dat' dataframe for prediction
  interval = "confidence", level = 0.95)$fit %>%  # Calculate 95% confidence intervals for the predictions
  as_tibble() %>%  # Convert the output to a tibble
  bind_cols(fict_dat, .)  # Combine the predictions with the original 'fict_dat' 

ggplot(data = fict_dat, aes(y = fit, x = years_indep)) +  # Create a plot using the predicted data
  geom_vline(xintercept = c(34, 236), color = "red", lty = "dashed") +  # Add two vertical dashed lines at x = 34 and x = 236
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.5) +  # Add a ribbon representing the confidence interval
  geom_line() +  # Add a line for the predicted values
  labs(
    title = "Predictions based on multiple OLS of extreme poverty",  # Set the plot title
    x = "By years since independence from colonialism\n(Colonializer set to: Denmark)",  # Set the x-axis label
    y = "Predicted average of extreme poverty",  # Set the y-axis label
    caption = "") +  # Remove the caption
  theme_minimal()  # Use a minimal theme
```
`r unhide()`
```{r ref.label = "predplot2", out.width='100%', fig.height = 4, fig.width = 6, echo = FALSE}
```