---
title: "Exercise 1 (20 minutes)"
subtitle: "When you're done, assist some of the others ..."
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, servr.interval = 0.5, width = 115, digits = 2)
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, 
  warning = FALSE, cache = FALSE,
  comment = "#", strip.white = TRUE)

library(webexercises)
library(essentials)
```

1. Let's take another look at our research question on the legacy of colonialism. Use the setup below.

```{r}
pacman::p_load(
  tidyverse,  # Load the tidyverse package for data manipulation and visualization
  ggplot2,  # Load the ggplot2 package for creating plots
  wbstats,  # Load the wbstats package to download data from the World Bank
  estimatr,  # Load the estimatr package for running regressions with robust standard errors
  modelsummary)  # Load the modelsummary package for creating nice-looking regression tables

# 1. The worldbank data
Dat <- wb_data("SI.POV.DDAY",  # Download data on poverty headcount ratio at $1.90 a day from the World Bank
              start_date = 2002, end_date = 2022) %>%  # Specify the time range for the data
  rename(poverty = SI.POV.DDAY) %>%  # Rename the 'SI.POV.DDAY' variable to 'poverty'
  select(country, date, poverty) %>%  # Keep only the 'country', 'date', and 'poverty' variables
  drop_na(poverty) %>%  # Remove rows with missing 'poverty' values
  group_by(country) %>%  # Group the data by country
  filter(date == max(date)) %>%  # Keep only the most recent observation for each country
  mutate(date = as.numeric(date)) %>%  # Convert the 'date' variable to numeric
  ungroup()  # Ungroup the data

# 2. Colonizer data
Dat <- Dat %>% 
  mutate(
    colonizer = case_when(  # Create a new variable 'colonizer' based on the country name
      str_detect(country, "Algeria|Benin|Burkina Faso|Cambodia|Central African Republic|Chad") |  # If the country name contains any of these strings...
        str_detect(country, "Djibouti|Gabon|Guinea|Haiti|Laos|Lebanon|Madagascar|Mali|Congo, Rep.|Cote D'Ivoire") | 
        str_detect(country, "Mauritania|Niger|Senegal|Syria|Togo|Tunisia|Vietnam") ~ "France",  # ...then assign "France" as the colonizer
      str_detect(country, "Angola|Brazil|Mozambique|Guinea-Bissau") ~ "Portugal",  # Similarly, assign "Portugal" for these countries
      str_detect(country, "Argentina|Bolivia|Chile|Colombia|Costa Rica|Cuba|Dominican Republic|Ecuador") |
        str_detect(country, "El Salvador|Guatemala|Honduras|Mexico|Nicaragua|Panama|Paraguay|Peru") |
        str_detect(country, "Uruguay|Venezuela") ~ "Spain",  # Assign "Spain" for these countries
      str_detect(country, "Australia|Bahrain|Bangladesh|Botswana|Canada|Cyprus|Egypt|Fiji") |
        str_detect(country, "Gambia|Ghana|Guyana|India|Iraq|Israel|Jamaica|Jordan|Kenya|Kuwait") |
        str_detect(country, "Lesotho|Malawi|Malaysia|Mauritius|Myanmar|New Zealand|Nigeria|Oman") |
        str_detect(country, "Pakistan|Qatar|Sierra Leone|Singapore|South Africa") |
        str_detect(country, "Sri Lanka|Sudan|Swaziland|Tanzania|Trinidad/Tobago|Uganda") |
        str_detect(country, "United Arab Emirates|United States|Yemen|Zambia|Zimbabwe") ~ "Britain",  # Assign "Britain" for these countries
      str_detect(country, "Burundi|Congo, Dem. Rep.|Rwanda") ~ "Belgium",  # Assign "Belgium" for these countries
      str_detect(country, "Iceland|Estonia") ~ "Denmark"))  # Assign "Denmark" for these countries

# 3. Independence from colonialism data
Dat <- Dat %>% 
  mutate(
    years_indep = case_when(  # Create a new variable 'years_indep' (years of independence) based on the country and colonization history
      is.na(colonizer) ~ as.numeric(NA),  # If the 'colonizer' is NA, then 'years_indep' is also NA
      country == "Estonia" ~ date - 1645,  # For Estonia, calculate years of independence since 1645
      country == "United States" ~ date - 1783, country == "Haiti" ~ date - 1804,  # Similar calculations for other countries based on their independence years
      country == "Paraguay" ~ date - 1811, country == "Chile" ~ date - 1818,
      str_detect(country, "Argentina|Bolivia|Colombia") ~ date - 1819,
      str_detect(country, "Costa Rica|Dominican Republic|Mexico|Nicaragua|Panama|El Salvador|Guatemala|Honduras|Venezuela") ~ date - 1821,
      str_detect(country, "Brazil|Ecuador") ~ date - 1822,
      country == "Peru" ~ date - 1824, country == "Uruguay" ~ date - 1828,
      country == "Cuba" ~ date - 1899, country == "Australia" ~ date - 1901,
      country == "New Zealand" ~ date - 1907, country == "South Africa" ~ date - 1910,
      country == "Egypt" ~ date - 1922, country == "Iraq" ~ date - 1932,
      str_detect(country, "Iceland") ~ date - 1944,
      str_detect(country, "Korea|Taiwan|Vietnam") ~ date - 1945,
      str_detect(country, "Lebanon|Philippines|Syria") ~ date - 1946,
      str_detect(country, "Bangladesh|Pakistan|India|Liberia") ~ date - 1947,
      str_detect(country, "Myanmar|Israel|Jordan|Sri Lanka") ~ date - 1948,
      country == "Indonesia" ~ date - 1949, country == "Libya" ~ date - 1951,
      str_detect(country, "Cambodia|Loas") ~ date - 1954,
      str_detect(country, "Morocco|Sudan|Tunisia") ~ date - 1956,
      str_detect(country, "Malaysia|Ghana") ~ date - 1957,
      country == "Guinea" ~ date - 1958, country == "Singapore" ~ date - 1959,
      str_detect(country, "Benin|Burkina Faso|Central African Republic|Chad|Congo, Dem. Rep.|Congo, Rep.|Code D'Ivoire|Mali|Mauritania|Niger|Nigeria|Senegal|Gabon|Somalia|Togo") ~ date - 1960,
      str_detect(country, "Kuwait|Sierra Leone|Tanzania") ~ date - 1961,
      str_detect(country, "Algeria|Burundi|Rwanda|Jamaica|Trinidad/ Tobago|Uganda") ~ date - 1962,
      country == "Kenya" ~ date - 1963,
      str_detect(country, "Malawi|Zambia") ~ date - 1964,
      str_detect(country, "Gambia|Zimbabwe") ~ date - 1965,
      str_detect(country, "Botswana|Lesotho|Guyana") ~ date - 1966,
      str_detect(country, "Canada|Yemen") ~ date - 1967,
      str_detect(country, "Mauritius|Swaziland") ~ date - 1968,
      country == "Fiji" ~ date - 1970,
      str_detect(country, "Bahrain|Oman|Qatar|United Arab Emirates") ~ date - 1971,
      country == "Guinea-Bissau" ~ date - 1974,
      str_detect(country, "Angola|Mozambigue|Papua New Guinea") ~ date - 1975,
      country == "Djibouti" ~ date - 1977, 
      country == "Namibia" ~ date - 1990))
```

2. Make Denmark the reference category. Ask an AI how you can use `fct_relevel("Denmark")` to achieve this.
`r hide("R solution -> dont' peak to early ;) !")`
```{r}
Dat <- Dat %>%
  mutate(  # Modify the 'colonizer' variable
    # Relevel the 'colonizer' factor so that "Denmark" is the reference category
    colonizer = colonizer %>% fct_relevel("Denmark"))  
```
`r unhide()`

3. Answer the following question: To which degree do differences in the number of years since independence from colonialism explain why we saw significant differences in poverty rates between countries colonized by different colonial powers? `r mcq(c("It does not explain much", "Many of the stars are gone.", answer = "The years since indepence explain a large share of the different poverty legacies of the various colonializers"))`. 

`r hide("R solution -> dont' peak to early ;) !")`
```{r}
# Run a robust bivariate linear regression of 'poverty' on 'colonizer' using the 'Dat' dataset. 
# This model examines the relationship between the colonizing country and poverty levels.
ols_bi <- lm_robust(poverty ~ colonizer, data = Dat)  

# Run a robust multiple linear regression of 'poverty' on 'colonizer' and 'years_indep' (years of independence). 
# This model adds years of independence as a control variable to assess the impact of the colonizer while accounting for the time since independence.
ols_mult <- lm_robust(poverty ~ colonizer + years_indep, data = Dat)  

# Create a summary table for both regression models ('ols_bi' and 'ols_mult'), labeling them "Biv. OLS" and "Mult. OLS" respectively.
modelsummary(list("Biv. OLS" = ols_bi, "Mult. OLS" = ols_mult),   
             gof_map = c("nobs", "r.squared"),  # Include the number of observations ('nobs') and R-squared ('r.squared').
             stars = TRUE)  # Add stars to indicate the statistical significance.
```
`r unhide()`

4. What does this say about a causal colonizer effect? `r mcq(c("It's much smaller than the bivariate result suggests.", "Correlation isn't causation", answer = "It's unclear whether the years since independence are not themselves an outcome of the colonializer (i.e. a mediator).", "It's larger than the bivariate result suggests."))`